---
title: "Phase 3a: Structural Topic Model for Discourse Framing"
subtitle: "STM preprocessing, K selection, topic diagnostics, frame mapping, prevalence estimation"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

# Design Notes

Phase 3a fits a Structural Topic Model (STM) to the analytical corpus to
discover latent discourse frames inductively. This complements the
dictionary-based register classification from Phase 2c by allowing the
data to surface frames the researcher did not anticipate.

Seven components are produced.

1. Text preprocessing: tokenisation, Croatian stopword removal,
   trimming, and DFM construction via quanteda.
2. K selection via held-out likelihood, semantic coherence, and
   exclusivity across a range of candidate K values.
3. Final model fit at the selected K with prevalence covariates
   (destination, quarter, platform category).
4. Topic diagnostics: top words by probability and FREX, topic
   proportions, coherence-exclusivity scatter.
5. Topic-to-frame labelling based on top words and exemplary documents.
6. Prevalence estimation via estimateEffect() to test how topic
   prevalence varies by destination, time, and platform.
7. Convergent validity: comparison of STM-derived frames with the
   dictionary-based register classification from Phase 2c.

The primary text field is MENTION_SNIPPET (the contextual window around
the destination mention). FULL_TEXT is available for robustness checks
but is noisier because it includes content unrelated to the destination.


## 0 Configuration

```{r}
#| label: config

config_path <- "data/config_phase3a.rds"

if (file.exists(config_path)) {
  config <- readRDS(config_path)
  cat("Phase 3a configuration loaded from disk.\n")
} else {
  config <- list(
    phase_name    = "03a_stm_framing",
    timestamp     = Sys.time(),
    seed          = 2026L,

    # -- Input paths --
    input_paths = list(
      corpus = "../02-descriptive/data/destinations_analytical_final.rds",
      corpus_fallback = "../02-descriptive/data/destinations_analytical_cleaned.rds"
    ),

    # -- Output paths --
    output_paths = list(
      dfm            = "data/03a_dfm.rds",
      searchK        = "data/03a_searchK.rds",
      stm_model      = "data/03a_stm_model.rds",
      stm_effect     = "data/03a_stm_effect.rds",
      topic_labels   = "data/03a_topic_labels.rds",
      topic_docs     = "data/03a_topic_exemplars.rds",
      convergence    = "data/03a_convergence_check.rds",
      config         = config_path,
      metadata       = "output/metadata/meta_03a_stm_framing.rds"
    ),

    # -- Preprocessing parameters --
    preprocessing = list(
      text_field        = "MENTION_SNIPPET",
      min_termfreq      = 10L,
      max_docfreq_pct   = 0.90,
      min_nchar         = 3L,
      ngram_range       = c(1L, 2L),
      extra_stopwords   = c(
        "http", "https", "www", "com", "hr", "foto", "video",
        "izvor", "objav", "članak", "preuz", "pročitajte", "više",
        "klik", "nastavak", "tekst", "autor", "urednik",
        "godina", "godine", "godinu", "godini", "dan", "dana",
        "biti", "jest", "može", "trebati", "imati", "ići",
        "sve", "svi", "koji", "koja", "koje", "kojeg", "kojoj",
        "jedan", "jedna", "jedno", "dva", "tri", "prvi", "drugi",
        "novi", "nova", "novo", "veliki", "velika", "veliko",
        "već", "još", "samo", "također", "tako", "isto",
        "danas", "jučer", "sutra", "sada", "ovdje", "gdje",
        "ovaj", "ovog", "ovom", "taj", "tog", "tom",
        "cijeli", "cijela", "cijelom", "svaki", "svaka", "svako",
        "share", "tweet", "plus", "ispis", "print",
        "facebook", "instagram", "twitter",
        "kolačić", "kolačiće", "kolačića",
        "kliknite", "messenger", "viber", "whatsapp",
        "preuzeto", "objavljeno", "objavio"
      )
    ),

    # -- K selection parameters --
    k_selection = list(
      k_range         = c(5L, 7L, 9L, 10L, 12L, 15L, 18L, 20L, 25L, 30L, 35L),
      heldout_prop    = 0.10,
      max_em_iter     = 150L
    ),

    # -- Final model parameters --
    model = list(
      K               = NULL,    # set after searchK
      prevalence_formula = "~ destination + dest_type_cat + s(year_num) + feast_window",
      max_em_iter     = 500L,
      convergence_tol = 1e-5,
      n_exemplars     = 15L
    ),

    # -- Dictionary paths --
    dictionary_paths = list(
      destinations = "../../dictionaries/destinations.csv",
      registers    = "../../dictionaries/registers.csv",
      stopwords_hr = NULL  # built-in quanteda stopwords("hr")
    )
  )
  saveRDS(config, config_path)
  cat("Phase 3a configuration saved.\n")
}
```


## 1 Setup and Load Corpus

```{r}
#| label: setup

library(data.table)
library(stringi)
library(quanteda)
library(quanteda.textstats)
library(stm)
library(ggplot2)
library(patchwork)
library(knitr)
library(digest)

source("../../R/theme_reltur.R")
source("../../R/decision_gate.R")
source("../../R/metadata_snapshot.R")
source("../../R/strip_boilerplate.R")
source("../../R/join_dest_type.R")

for (d in c("data", "data-raw", "output/tables", "output/figures",
            "output/reports", "output/metadata")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}

set.seed(config$seed)
```

```{r}
#| label: load-corpus

# Try loading in order of preference
corpus_loaded <- FALSE
for (p in config$input_paths) {
  if (file.exists(p)) {
    dt <- readRDS(p)
    setDT(dt)
    cat(sprintf("Loaded corpus from: %s\n", p))
    cat(sprintf("  Rows: %s | Columns: %s\n", format(nrow(dt), big.mark = " "),
                ncol(dt)))
    corpus_loaded <- TRUE
    break
  }
}
stopifnot(corpus_loaded)
dt <- dt[cleaning_action == "keep"]
cat(sprintf("  After cleaning filter: %s rows\n", format(nrow(dt), big.mark = " ")))

dt <- strip_boilerplate(dt)
dt <- join_dest_type(dt)
dt[, year_num := as.integer(format(date, "%Y")) - 2021L]
if ("feast_name" %in% names(dt)) {
  dt[, feast_window := as.integer(!is.na(feast_name))]
} else if (!"feast_window" %in% names(dt)) {
  dt[, feast_window := 0L]
  cat("Note: feast_name not found; feast_window set to 0 for all rows.\n")
}

if (!"yearmonth" %in% names(dt))
  dt[, yearmonth := as.Date(format(date, "%Y-%m-01"))]

# Verify text field exists and is non-empty
text_field <- config$preprocessing$text_field
stopifnot(text_field %in% names(dt))

n_missing_text <- sum(is.na(dt[[text_field]]) | dt[[text_field]] == "")
cat(sprintf("Missing/empty %s: %d (%.1f%%)\n",
            text_field, n_missing_text,
            100 * n_missing_text / nrow(dt)))
```


## 2 Text Preprocessing

```{r}
#| label: preprocess-text

# ---- 2.1 Build working copy with clean text ----
stm_dt <- dt[!is.na(get(text_field)) & get(text_field) != ""]

# Lowercase with Croatian locale
stm_dt[, text_clean := stri_trans_tolower(get(text_field), locale = "hr")]

# Remove URLs
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "https?://\\S+", " ")]

# Remove email addresses
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "\\S+@\\S+\\.\\S+", " ")]

# Remove HTML entities and tags
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "<[^>]+>", " ")]
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "&\\w+;", " ")]

# Remove digits (but keep words containing digits like "covid19")
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "\\b\\d+\\b", " ")]

# Normalise whitespace
stm_dt[, text_clean := stri_replace_all_regex(text_clean, "\\s+", " ")]
stm_dt[, text_clean := stri_trim_both(text_clean)]

# Drop documents that became empty after cleaning
n_before <- nrow(stm_dt)
stm_dt <- stm_dt[stri_length(text_clean) >= 20L]
cat(sprintf("Documents after text cleaning: %d (dropped %d short/empty)\n",
            nrow(stm_dt), n_before - nrow(stm_dt)))
```

```{r}
#| label: build-dfm

# ---- 2.2 Croatian stopwords ----
stopwords_hr <- quanteda::stopwords("hr", source = "stopwords-iso")
extra_sw     <- config$preprocessing$extra_stopwords
all_stopwords <- unique(c(stopwords_hr, extra_sw))
cat(sprintf("Total stopwords: %d (ISO: %d + project: %d)\n",
            length(all_stopwords), length(stopwords_hr), length(extra_sw)))

# ---- 2.3 Quanteda corpus and tokenisation ----
qcorp <- quanteda::corpus(stm_dt$text_clean,
                          docvars = stm_dt[, .(pair_id, destination,
                                               dest_type_cat, year_num,
                                               feast_window, date, yearmonth,
                                               has_religious_register,
                                               has_tourism_register)])

qtok <- quanteda::tokens(qcorp,
                         remove_punct  = TRUE,
                         remove_symbols = TRUE,
                         remove_numbers = TRUE,
                         remove_url     = TRUE) |>
  quanteda::tokens_remove(all_stopwords, padding = FALSE) |>
  quanteda::tokens_select(min_nchar = config$preprocessing$min_nchar)

# ---- 2.4 Ngrams (unigrams + bigrams) ----
if (config$preprocessing$ngram_range[2] > 1L) {
  qtok_ngram <- quanteda::tokens_ngrams(qtok,
                                        n = config$preprocessing$ngram_range[1]:
                                            config$preprocessing$ngram_range[2])
  # Combine unigrams and bigrams
  qtok <- quanteda::tokens_compound(qtok,
            quanteda::phrase(quanteda::featnames(
              quanteda::dfm(qtok_ngram) |>
                quanteda::dfm_trim(min_termfreq = config$preprocessing$min_termfreq * 2L)
            )))
}

# ---- 2.5 Document-feature matrix ----
raw_dfm <- quanteda::dfm(qtok)
cat(sprintf("Raw DFM: %d docs x %d features\n", nrow(raw_dfm), ncol(raw_dfm)))

# Trim rare and overly common terms
trimmed_dfm <- quanteda::dfm_trim(
  raw_dfm,
  min_termfreq  = config$preprocessing$min_termfreq,
  max_docfreq   = config$preprocessing$max_docfreq_pct,
  docfreq_type  = "prop"
)
cat(sprintf("Trimmed DFM: %d docs x %d features\n",
            nrow(trimmed_dfm), ncol(trimmed_dfm)))

# Remove documents that became empty after trimming
keep_docs <- quanteda::rowSums(trimmed_dfm) > 0
trimmed_dfm <- trimmed_dfm[keep_docs, ]
cat(sprintf("After removing empty docs: %d docs x %d features\n",
            nrow(trimmed_dfm), ncol(trimmed_dfm)))
```

```{r}
#| label: top-features

# ---- 2.6 Inspect top features ----
top50 <- quanteda::topfeatures(trimmed_dfm, n = 50)
top_dt <- data.table(feature = names(top50), frequency = unname(top50))
knitr::kable(top_dt, caption = "Top 50 features after preprocessing")
```

```{r}
#| label: convert-stm

# ---- 2.7 Convert to STM format ----
# Build covariates for prevalence model
doc_meta <- quanteda::docvars(trimmed_dfm)
doc_meta <- as.data.table(doc_meta)

# Quarter as numeric for spline (months since start)
doc_meta[, quarter_num := as.numeric(difftime(yearmonth,
           min(yearmonth, na.rm = TRUE), units = "days")) / 30.44]

# Platform category (simplified)
if ("SOURCE_TYPE" %in% names(stm_dt)) {
  doc_meta[, platform_cat := stm_dt[keep_docs][["SOURCE_TYPE"]]]
} else {
  doc_meta[, platform_cat := "web"]
}
doc_meta[, platform_cat := fifelse(
  platform_cat %in% c("web", "portal", "online"), "web",
  fifelse(platform_cat %in% c("facebook", "twitter", "instagram",
                               "reddit", "youtube"), "social",
          "other"))]
doc_meta[, platform_cat := as.factor(platform_cat)]
doc_meta[, destination := as.factor(destination)]

# New covariates from 02c enrichment
if (!"dest_type_cat" %in% names(doc_meta)) {
  doc_meta[, dest_type_cat := stm_dt[keep_docs][["dest_type_cat"]]]
}
doc_meta[, dest_type_cat := as.factor(dest_type_cat)]
if (!"year_num" %in% names(doc_meta)) {
  doc_meta[, year_num := stm_dt[keep_docs][["year_num"]]]
}
if (!"feast_window" %in% names(doc_meta)) {
  doc_meta[, feast_window := stm_dt[keep_docs][["feast_window"]]]
}
doc_meta[is.na(feast_window), feast_window := 0L]

stm_input <- quanteda::convert(trimmed_dfm, to = "stm")
stm_input$meta <- as.data.frame(doc_meta)

saveRDS(list(dfm = trimmed_dfm, stm_input = stm_input, doc_meta = doc_meta),
        config$output_paths$dfm)
cat("DFM and STM input saved.\n")
```


## 3 K Selection (searchK)

```{r}
#| label: searchK
#| cache: true

set.seed(config$seed)

cat(sprintf("Running searchK for K = %s\n",
            paste(config$k_selection$k_range, collapse = ", ")))
cat("This may take 30-60 minutes depending on corpus size.\n")

searchK_out <- stm::searchK(
  documents  = stm_input$documents,
  vocab      = stm_input$vocab,
  K          = config$k_selection$k_range,
  prevalence  = ~ destination + dest_type_cat + s(year_num) + feast_window,
  data        = stm_input$meta,
  heldout.seed = config$seed,
  proportion   = config$k_selection$heldout_prop,
  max.em.its   = config$k_selection$max_em_iter,
  init.type    = "Spectral",
  seed         = config$seed,
  verbose      = TRUE
)

saveRDS(searchK_out, config$output_paths$searchK)
cat("searchK results saved.\n")
```

```{r}
#| label: plot-searchK

# ---- 3.1 Extract and plot diagnostics ----
# searchK returns list-columns; unlist them for plotting
safe_unlist_numeric <- function(x) {
  v <- unlist(x)
  if (is.null(v)) return(rep(NA_real_, length(unlist(searchK_out$results$K))))
  as.numeric(v)
}

sk <- data.table(
  K        = as.integer(unlist(searchK_out$results$K)),
  heldout  = safe_unlist_numeric(searchK_out$results$heldout),
  semcoh   = safe_unlist_numeric(searchK_out$results$semcoh),
  exclus   = safe_unlist_numeric(searchK_out$results$exclus),
  residual = safe_unlist_numeric(searchK_out$results$residual)
)

# Held-out likelihood
p_heldout <- ggplot(sk, aes(K, heldout)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  labs(title = "Held-out Likelihood", x = "K", y = "Log likelihood") +
  theme_reltur()

# Semantic coherence
p_coherence <- ggplot(sk, aes(K, semcoh)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  labs(title = "Semantic Coherence", x = "K", y = "Coherence") +
  theme_reltur()

# Exclusivity (may be NA if searchK did not compute it)
if (all(is.na(sk$exclus))) {
  p_exclusivity <- ggplot(sk, aes(K, semcoh)) +
    annotate("text", x = mean(sk$K), y = 0, label = "Exclusivity not available", size = 4) +
    labs(title = "Exclusivity", x = "K", y = "") +
    theme_reltur()
} else {
  p_exclusivity <- ggplot(sk, aes(K, exclus)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2.5) +
    labs(title = "Exclusivity", x = "K", y = "Exclusivity") +
    theme_reltur()
}

# Residuals
p_residual <- ggplot(sk, aes(K, residual)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  labs(title = "Residual Dispersion", x = "K", y = "Residuals") +
  theme_reltur()

p_searchK <- (p_heldout | p_coherence) / (p_exclusivity | p_residual) +
  patchwork::plot_annotation(
    title = "STM Model Selection Diagnostics",
    subtitle = sprintf("Candidate K = {%s}",
                        paste(config$k_selection$k_range, collapse = ", "))
  )

ggsave("output/figures/03a_searchK_diagnostics.png", p_searchK,
       width = 12, height = 9, dpi = 300, bg = "white")
p_searchK
```

```{r}
#| label: select-K

# ---- 3.2 Select K ----
# Select K based on held-out likelihood peak and coherence-exclusivity trade-off
# User should inspect the plots and may override this selection
best_idx <- which.max(sk$heldout)
selected_K <- sk$K[best_idx]
cat(sprintf("\nAutomatic K selection (max held-out): K = %d\n", selected_K))
cat("Review the diagnostic plots above. Override config$model$K if needed.\n")

# Update config
config$model$K <- selected_K
saveRDS(config, config$output_paths$config)
```


## 4 Fit Final Model

```{r}
#| label: fit-stm
#| cache: true

set.seed(config$seed)

K_final <- config$model$K
cat(sprintf("Fitting final STM with K = %d\n", K_final))

stm_fit <- stm::stm(
  documents    = stm_input$documents,
  vocab        = stm_input$vocab,
  K            = K_final,
  prevalence   = ~ destination + dest_type_cat + s(year_num) + feast_window,
  data         = stm_input$meta,
  max.em.its   = config$model$max_em_iter,
  emtol        = config$model$convergence_tol,
  init.type    = "Spectral",
  seed         = config$seed,
  verbose      = TRUE
)

saveRDS(stm_fit, config$output_paths$stm_model)
cat("STM model saved.\n")

# Check convergence
cat(sprintf("Converged: %s (iterations: %d)\n",
            stm_fit$convergence$converged,
            length(stm_fit$convergence$bound)))
```


## 5 Topic Diagnostics

```{r}
#| label: topic-top-words

# ---- 5.1 Top words by Prob and FREX ----
top_words_prob <- stm::labelTopics(stm_fit, n = 15)

# Build tidy table
topics_dt <- rbindlist(lapply(1:K_final, function(k) {
  data.table(
    topic    = k,
    rank     = 1:15,
    prob     = top_words_prob$prob[k, ],
    frex     = top_words_prob$frex[k, ],
    lift     = top_words_prob$lift[k, ],
    score    = top_words_prob$score[k, ]
  )
}))

# Display concise version (top 10)
topics_wide <- dcast(
  topics_dt[rank <= 10],
  topic ~ rank,
  value.var = "frex"
)
setnames(topics_wide, c("Topic", paste0("FREX_", 1:10)))
knitr::kable(topics_wide, caption = "Top 10 FREX words per topic")
```

```{r}
#| label: topic-top-words-prob

topics_wide_prob <- dcast(
  topics_dt[rank <= 10],
  topic ~ rank,
  value.var = "prob"
)
setnames(topics_wide_prob, c("Topic", paste0("Prob_", 1:10)))
knitr::kable(topics_wide_prob, caption = "Top 10 Highest Probability words per topic")
```

```{r}
#| label: topic-proportions

# ---- 5.2 Topic proportions ----
theta <- stm_fit$theta  # doc x topic matrix
topic_means <- colMeans(theta)
topic_sds   <- apply(theta, 2, sd)

prop_dt <- data.table(
  topic      = 1:K_final,
  mean_prop  = topic_means,
  sd_prop    = topic_sds,
  median_prop = apply(theta, 2, median),
  max_prop   = apply(theta, 2, max)
)

# Flag potentially problematic topics
prop_dt[, flag := fifelse(mean_prop < 0.02, "SMALL",
                   fifelse(mean_prop > 0.25, "BROAD", "OK"))]

knitr::kable(prop_dt[order(-mean_prop)], digits = 4,
             caption = "Topic proportion summary")

# Plot
p_prop <- ggplot(prop_dt, aes(x = reorder(factor(topic), -mean_prop),
                               y = mean_prop)) +
  geom_col(fill = "#4A90D9", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_prop - sd_prop,
                    ymax = mean_prop + sd_prop),
                width = 0.3, linewidth = 0.4) +
  geom_hline(yintercept = 1 / K_final, linetype = "dashed", color = "red") +
  labs(title = sprintf("Topic Proportions (K = %d)", K_final),
       subtitle = "Red line = uniform distribution; bars = mean +/- SD",
       x = "Topic", y = "Mean proportion") +
  theme_reltur()

ggsave("output/figures/03a_topic_proportions.png", p_prop,
       width = 10, height = 6, dpi = 300, bg = "white")
p_prop
```

```{r}
#| label: coherence-exclusivity

# ---- 5.3 Coherence vs Exclusivity scatter ----
coh <- stm::semanticCoherence(stm_fit, stm_input$documents)
exc <- stm::exclusivity(stm_fit)

ce_dt <- data.table(topic = 1:K_final, coherence = coh, exclusivity = exc)

p_ce <- ggplot(ce_dt, aes(coherence, exclusivity, label = topic)) +
  geom_point(size = 3, color = "#4A90D9") +
  geom_text(vjust = -0.8, size = 3.5) +
  labs(title = "Semantic Coherence vs Exclusivity",
       subtitle = sprintf("K = %d | Top-right = ideal topics", K_final),
       x = "Semantic Coherence", y = "Exclusivity") +
  theme_reltur()

ggsave("output/figures/03a_coherence_exclusivity.png", p_ce,
       width = 8, height = 6, dpi = 300, bg = "white")
p_ce
```

```{r}
#| label: topic-correlation

# ---- 5.4 Topic correlation network ----
topic_corr <- stm::topicCorr(stm_fit, method = "simple", verbose = FALSE)

corr_mat <- topic_corr$cor
corr_dt <- as.data.table(as.table(corr_mat))
setnames(corr_dt, c("topic_a", "topic_b", "correlation"))
corr_dt[, topic_a := as.integer(topic_a)]
corr_dt[, topic_b := as.integer(topic_b)]
corr_dt <- corr_dt[topic_a < topic_b & abs(correlation) > 0.05]

if (nrow(corr_dt) > 0) {
  knitr::kable(corr_dt[order(-abs(correlation))][1:min(20, .N)],
               digits = 3,
               caption = "Top topic-topic correlations (|r| > 0.05)")
} else {
  cat("No strong topic correlations detected.\n")
}
```


## 6 Exemplary Documents

```{r}
#| label: exemplars

# ---- 6.1 Find exemplary documents per topic ----
n_exemplars <- config$model$n_exemplars

exemplar_list <- lapply(1:K_final, function(k) {
  # Documents with highest theta for this topic
  top_idx <- order(theta[, k], decreasing = TRUE)[1:n_exemplars]

  data.table(
    topic    = k,
    rank     = 1:n_exemplars,
    doc_idx  = top_idx,
    theta_k  = theta[top_idx, k],
    pair_id  = stm_input$meta$pair_id[top_idx],
    destination = stm_input$meta$destination[top_idx],
    text     = stm_dt[keep_docs][top_idx][[text_field]]
  )
})

exemplars_dt <- rbindlist(exemplar_list)
saveRDS(exemplars_dt, config$output_paths$topic_docs)

# Show top 3 per topic (truncated text)
exemplars_dt[rank <= 3, .(topic, rank, destination,
  theta_k = round(theta_k, 3),
  text_snippet = stri_sub(text, 1, 120))] |>
  knitr::kable(caption = "Top 3 exemplary documents per topic (truncated)")
```


## 7 Topic-Frame Labelling

```{r}
#| label: topic-labels

# ---- 7.1 Automated label suggestion ----
# Combine FREX words + destination affinity to suggest labels
# This is a starting point; manual refinement is expected

dest_topic_prop <- as.data.table(stm_input$meta)[, destination := as.character(destination)]
dest_topic_prop <- cbind(dest_topic_prop, as.data.table(theta))

# Mean theta per destination per topic
dest_aff <- dest_topic_prop[, lapply(.SD, mean),
  by = destination,
  .SDcols = paste0("V", 1:K_final)]

dest_aff_long <- melt(dest_aff, id.vars = "destination",
                       variable.name = "topic", value.name = "mean_theta")
dest_aff_long[, topic := as.integer(stri_replace_first_fixed(topic, "V", ""))]

# For each topic, which destinations have above-average prevalence?
dest_aff_long[, overall_mean := mean(mean_theta), by = topic]
dest_aff_long[, above_avg := mean_theta > overall_mean * 1.2]

top_dests_per_topic <- dest_aff_long[above_avg == TRUE,
  .(top_destinations = paste(destination, collapse = ", ")),
  by = topic]

# Combine with top words for label suggestion
label_dt <- merge(
  topics_dt[rank <= 5, .(top_frex = paste(frex, collapse = ", ")), by = topic],
  top_dests_per_topic,
  by = "topic", all.x = TRUE
)

# Register affinity: what share of each topic's documents have religious/tourism register?
register_aff <- dest_topic_prop[, {
  reg_rel <- has_religious_register
  reg_tour <- has_tourism_register
  lapply(1:K_final, function(k) {
    w <- get(paste0("V", k))
    list(
      pct_religious = weighted.mean(reg_rel, w, na.rm = TRUE),
      pct_tourism   = weighted.mean(reg_tour, w, na.rm = TRUE)
    )
  })
}]

# Simpler approach: per-document dominant topic, then register percentages
dom_topic <- apply(theta, 1, which.max)
reg_dt <- data.table(
  dom_topic = dom_topic,
  has_rel   = stm_input$meta$has_religious_register,
  has_tour  = stm_input$meta$has_tourism_register
)
reg_summary <- reg_dt[, .(
  n         = .N,
  pct_rel   = 100 * mean(has_rel, na.rm = TRUE),
  pct_tour  = 100 * mean(has_tour, na.rm = TRUE)
), by = dom_topic]
setnames(reg_summary, "dom_topic", "topic")

label_dt <- merge(label_dt, reg_summary, by = "topic", all.x = TRUE)

# Suggest frame category based on register prevalence
label_dt[, suggested_frame := fifelse(
  pct_rel > 60 & pct_tour < 30, "religijski",
  fifelse(pct_tour > 60 & pct_rel < 30, "turistički",
  fifelse(pct_rel > 40 & pct_tour > 40, "mješoviti_religijsko-turistički",
  fifelse(pct_rel < 30 & pct_tour < 30, "opći/politički",
          "neodređeno"))))]

knitr::kable(label_dt, digits = 1,
             caption = "Topic label suggestions (automated; requires manual review)")

saveRDS(label_dt, config$output_paths$topic_labels)
```

```{r}
#| label: dest-topic-heatmap

# ---- 7.2 Destination × Topic heatmap ----
dest_aff_wide <- dcast(dest_aff_long, destination ~ topic, value.var = "mean_theta")

# Normalise per topic for better visualisation
dest_aff_norm <- melt(dest_aff_long[, .(destination, topic,
  norm_theta = (mean_theta - min(mean_theta)) /
               (max(mean_theta) - min(mean_theta) + 1e-10)),
  by = topic], id.vars = c("destination", "topic"),
  measure.vars = "norm_theta")

p_heatmap <- ggplot(dest_aff_long,
       aes(x = factor(topic), y = destination, fill = mean_theta)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(aes(label = sprintf("%.2f", mean_theta)), size = 2.8) +
  scale_fill_gradient2(low = "white", mid = "#C6DBEF", high = "#08519C",
                       midpoint = 1 / K_final,
                       name = "Mean \u03B8") +
  labs(title = "Topic Prevalence by Destination",
       subtitle = "Mean document-topic proportion per destination",
       x = "Topic", y = NULL) +
  theme_reltur() +
  theme(axis.text.y = element_text(size = 10))

ggsave("output/figures/03a_dest_topic_heatmap.png", p_heatmap,
       width = max(10, K_final * 0.8), height = 7, dpi = 300, bg = "white")
p_heatmap
```


## 8 Prevalence Estimation

```{r}
#| label: estimate-effect
#| cache: true

# ---- 8.1 estimateEffect for destination and time ----
set.seed(config$seed)

effect_fit <- stm::estimateEffect(
  formula   = 1:K_final ~ destination + dest_type_cat + s(year_num) + feast_window,
  stmobj    = stm_fit,
  metadata  = stm_input$meta,
  uncertainty = "Global"
)

saveRDS(effect_fit, config$output_paths$stm_effect)
cat("estimateEffect saved.\n")

# Feast window effect
effect_feast <- stm::estimateEffect(
  formula     = 1:K_final ~ feast_window,
  stmobj      = stm_fit,
  metadata    = stm_input$meta,
  uncertainty = "Global"
)
saveRDS(effect_feast, "data/03a_stm_effect_feast.rds")

# Year trend effect
effect_year <- stm::estimateEffect(
  formula     = 1:K_final ~ s(year_num),
  stmobj      = stm_fit,
  metadata    = stm_input$meta,
  uncertainty = "Global"
)
saveRDS(effect_year, "data/03a_stm_effect_year.rds")
cat("Feast and year effects saved.\n")
```

```{r}
#| label: plot-destination-effects

# ---- 8.2 Destination effects for each topic ----
# Extract destination coefficients
dest_effects <- lapply(1:K_final, function(k) {
  summ <- summary(effect_fit, topics = k)
  coef_table <- summ$tables[[1]]
  coef_dt <- as.data.table(coef_table, keep.rownames = "term")
  coef_dt[, topic := k]
  coef_dt
})
dest_effects_dt <- rbindlist(dest_effects)
setnames(dest_effects_dt, c("term", "estimate", "se", "t_val", "p_val", "topic"))

# Filter to destination terms only
dest_coefs <- dest_effects_dt[stri_detect_regex(term, "^destination")]
dest_coefs[, destination := stri_replace_first_fixed(term, "destination", "")]

# Significance markers
dest_coefs[, sig := fifelse(p_val < 0.001, "***",
                    fifelse(p_val < 0.01, "**",
                    fifelse(p_val < 0.05, "*", "")))]

# Save table
fwrite(dest_coefs, "output/tables/03a_destination_effects.csv")

# Plot top topics (those with most destination variation)
dest_var <- dest_coefs[, .(dest_range = max(estimate) - min(estimate)), by = topic]
top_varying <- dest_var[order(-dest_range)][1:min(6, K_final)]$topic

p_dest_list <- lapply(top_varying, function(k) {
  plot_dt <- dest_coefs[topic == k]
  ggplot(plot_dt, aes(x = reorder(destination, estimate), y = estimate)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_pointrange(aes(ymin = estimate - 1.96 * se,
                        ymax = estimate + 1.96 * se),
                    color = "#4A90D9", size = 0.5) +
    coord_flip() +
    labs(title = sprintf("Topic %d", k),
         x = NULL, y = "Coefficient") +
    theme_reltur(base_size = 9)
})

p_dest_effects <- patchwork::wrap_plots(p_dest_list, ncol = 3) +
  patchwork::plot_annotation(
    title = "Destination Effects on Topic Prevalence",
    subtitle = "Coefficients with 95% CI (reference: aljmas)"
  )

ggsave("output/figures/03a_destination_effects.png", p_dest_effects,
       width = 14, height = max(6, length(top_varying) * 2), dpi = 300, bg = "white")
p_dest_effects
```

```{r}
#| label: plot-feast-effects

# ---- 8.3 Feast window effects ----
feast_effects <- rbindlist(lapply(seq_len(K_final), function(k) {
  e <- summary(effect_feast)$tables[[k]]
  data.table(
    topic = k,
    est   = e["feast_window", "Estimate"],
    se    = e["feast_window", "Std. Error"]
  )
}))
feast_effects[, sig := abs(est) > 1.96 * se]

p_feast <- ggplot(feast_effects,
    aes(x = reorder(factor(topic), est), y = est, colour = sig)) +
  geom_point() +
  geom_errorbar(aes(ymin = est - 1.96 * se, ymax = est + 1.96 * se),
                width = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  scale_colour_manual(values = c("TRUE" = "#B2182B", "FALSE" = "grey60")) +
  labs(x = "Topic", y = "Feast window effect (\u03B8 change)",
       colour = "Significant",
       title = "Feast Window Effect on Topic Prevalence") +
  theme_reltur()

ggsave("output/figures/03a_feast_effects.png", p_feast,
       width = 14, height = 14, dpi = 300, bg = "white")
p_feast
```

```{r}
#| label: plot-temporal-effects

# ---- 8.4 Temporal trends for selected topics ----
p_time_list <- lapply(1:min(K_final, 12), function(k) {
  stm::plot.estimateEffect(effect_year, covariate = "year_num",
                           topics = k, method = "continuous",
                           model = stm_fit,
                           main = sprintf("Topic %d", k),
                           xlab = "Years since 2021",
                           ylab = "Topic proportion")
})
# Note: stm::plot.estimateEffect writes to current device;
# for publication figures we extract the data manually

cat("Temporal effect plots generated (see HTML output).\n")
```


## 9 Convergent Validity: STM vs Register-Based Frames

```{r}
#| label: convergent-validity

# ---- 9.1 Map STM topics to register-based frames ----
# Dominant topic per document
dom_topic <- apply(theta, 1, which.max)

# Merge with register-based classification
valid_dt <- data.table(
  pair_id   = stm_input$meta$pair_id,
  stm_topic = dom_topic,
  has_rel   = stm_input$meta$has_religious_register,
  has_tour  = stm_input$meta$has_tourism_register
)

# Reconstruct register-based frame
valid_dt[, register_frame := fifelse(
  has_rel & !has_tour, "religious_only",
  fifelse(!has_rel & has_tour, "tourism_only",
  fifelse(has_rel & has_tour, "religious_tourism_mixed",
          "neither")))]

# Cross-tabulation: STM topic × register frame
cross_tab <- valid_dt[, .N, by = .(stm_topic, register_frame)]
cross_wide <- dcast(cross_tab, stm_topic ~ register_frame, value.var = "N", fill = 0)

knitr::kable(cross_wide,
             caption = "Cross-tabulation: STM dominant topic vs register-based frame")

fwrite(cross_wide, "output/tables/03a_stm_vs_register_crosstab.csv")

# ---- 9.2 Agreement metrics ----
# For each STM topic, what is its modal register frame?
topic_modal <- valid_dt[, .(
  n = .N,
  modal_frame = register_frame[which.max(tabulate(match(register_frame,
                  unique(register_frame))))],
  purity = max(tabulate(match(register_frame, unique(register_frame)))) / .N
), by = stm_topic][order(stm_topic)]

knitr::kable(topic_modal, digits = 3,
             caption = "STM topic purity against register-based frames")

# Overall agreement: % of documents where STM modal-frame matches register frame
valid_dt <- merge(valid_dt, topic_modal[, .(stm_topic, modal_frame)],
                  by = "stm_topic")
valid_dt[, agree := register_frame == modal_frame]
agreement_pct <- 100 * mean(valid_dt$agree, na.rm = TRUE)
cat(sprintf("\nOverall agreement (STM dominant topic → register frame): %.1f%%\n",
            agreement_pct))

# ---- 9.3 Adjusted Rand Index ----
if (requireNamespace("mclust", quietly = TRUE)) {
  ari <- mclust::adjustedRandIndex(valid_dt$stm_topic, valid_dt$register_frame)
  cat(sprintf("Adjusted Rand Index: %.3f\n", ari))
} else {
  ari <- NA_real_
  cat("mclust not available; skipping ARI.\n")
}

convergence_results <- list(
  cross_tab     = cross_wide,
  topic_purity  = topic_modal,
  agreement_pct = agreement_pct,
  ari           = ari
)
saveRDS(convergence_results, config$output_paths$convergence)
```

```{r}
#| label: convergence-heatmap

# ---- 9.4 Visualise convergence ----
conv_long <- melt(cross_wide, id.vars = "stm_topic",
                  variable.name = "register_frame", value.name = "count")

# Normalise within each topic (row-wise)
conv_long[, topic_total := sum(count), by = stm_topic]
conv_long[, pct := 100 * count / topic_total]

p_conv <- ggplot(conv_long, aes(x = register_frame, y = factor(stm_topic),
                                 fill = pct)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(aes(label = sprintf("%.0f%%", pct)), size = 3) +
  scale_fill_gradient(low = "white", high = "#2171B5", name = "% of topic") +
  labs(title = "STM Topic vs Register-Based Frame",
       subtitle = "Row-normalised: distribution of register frames within each STM topic",
       x = "Register-based frame", y = "STM Topic") +
  theme_reltur() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

ggsave("output/figures/03a_convergence_heatmap.png", p_conv,
       width = 10, height = max(6, K_final * 0.5), dpi = 300, bg = "white")
p_conv
```


## 10 Quality Profile and Summary

```{r}
#| label: summary-table

# ---- 10.1 Consolidated topic summary ----
summary_dt <- merge(prop_dt, ce_dt, by = "topic")
summary_dt <- merge(summary_dt, label_dt[, .(topic, top_frex, suggested_frame,
                                              pct_rel, pct_tour, n)],
                    by = "topic", all.x = TRUE)
summary_dt <- merge(summary_dt, topic_modal[, .(stm_topic, modal_frame, purity)],
                    by.x = "topic", by.y = "stm_topic", all.x = TRUE)

# Save
fwrite(summary_dt, "output/tables/03a_topic_summary.csv")
knitr::kable(summary_dt[order(-mean_prop)],
             digits = 3,
             caption = "Consolidated topic summary: proportion, coherence, exclusivity, register affinity, purity")
```

```{r}
#| label: save-tables

# ---- 10.2 Save all output tables ----
fwrite(topics_dt, "output/tables/03a_topic_words.csv")
fwrite(prop_dt, "output/tables/03a_topic_proportions.csv")
fwrite(ce_dt, "output/tables/03a_coherence_exclusivity.csv")
fwrite(dest_aff_long, "output/tables/03a_destination_topic_affinity.csv")

cat("All output tables saved to output/tables/.\n")
```


## 11 Decision Gate

```{r}
#| label: gate

gate_checks <- list(

  convergence = function() {
    if (stm_fit$convergence$converged) {
      list(status = "PASS",
           message = sprintf("Model converged in %d iterations",
                             length(stm_fit$convergence$bound)))
    } else {
      list(status = "CAUTION",
           message = sprintf("Model did NOT converge after %d iterations",
                             config$model$max_em_iter))
    }
  },

  no_degenerate_topics = function() {
    tiny <- sum(prop_dt$mean_prop < 0.01)
    if (tiny == 0) {
      list(status = "PASS", message = "No topics below 1% mean proportion")
    } else {
      list(status = "CAUTION",
           message = sprintf("%d topics below 1%% mean proportion", tiny))
    }
  },

  no_dominant_topics = function() {
    huge <- sum(prop_dt$mean_prop > 0.30)
    if (huge == 0) {
      list(status = "PASS", message = "No topics above 30% mean proportion")
    } else {
      list(status = "CAUTION",
           message = sprintf("%d topics above 30%% mean proportion", huge))
    }
  },

  min_purity = function() {
    min_pur <- min(topic_modal$purity, na.rm = TRUE)
    if (min_pur >= 0.40) {
      list(status = "PASS",
           message = sprintf("Minimum topic purity: %.1f%%", 100 * min_pur))
    } else {
      list(status = "CAUTION",
           message = sprintf("Low purity topic detected: %.1f%%", 100 * min_pur))
    }
  },

  register_agreement = function() {
    if (agreement_pct >= 50) {
      list(status = "PASS",
           message = sprintf("STM-register agreement: %.1f%%", agreement_pct))
    } else {
      list(status = "CAUTION",
           message = sprintf("Low STM-register agreement: %.1f%%", agreement_pct))
    }
  },

  corpus_coverage = function() {
    pct_used <- 100 * nrow(trimmed_dfm) / nrow(dt)
    if (pct_used >= 80) {
      list(status = "PASS",
           message = sprintf("%.1f%% of corpus retained after preprocessing", pct_used))
    } else {
      list(status = "CAUTION",
           message = sprintf("Only %.1f%% of corpus retained", pct_used))
    }
  }
)

gate_results <- run_decision_gate(gate_checks)
```


## 12 Metadata Snapshot

```{r}
#| label: metadata

output_files <- c(
  config$output_paths$dfm,
  config$output_paths$stm_model,
  config$output_paths$stm_effect,
  config$output_paths$topic_labels,
  config$output_paths$topic_docs,
  config$output_paths$convergence,
  "output/tables/03a_topic_summary.csv",
  "output/tables/03a_topic_words.csv",
  "output/tables/03a_destination_effects.csv",
  "output/figures/03a_searchK_diagnostics.png",
  "output/figures/03a_topic_proportions.png",
  "output/figures/03a_coherence_exclusivity.png",
  "output/figures/03a_dest_topic_heatmap.png",
  "output/figures/03a_destination_effects.png",
  "output/figures/03a_convergence_heatmap.png"
)

# Only checksum files that exist
output_files <- output_files[file.exists(output_files)]

metadata_snapshot(
  phase_name       = config$phase_name,
  config           = config,
  gate_results     = gate_results,
  output_files     = output_files,
  dictionary_files = unlist(config$dictionary_paths[!sapply(config$dictionary_paths, is.null)]),
  notes = sprintf(
    "STM K=%d fitted on %s field. %d docs, %d features. Agreement with register frames: %.1f%%.",
    K_final, config$preprocessing$text_field,
    nrow(trimmed_dfm), ncol(trimmed_dfm), agreement_pct
  )
)

cat("\n=== Phase 3a complete ===\n")
```
