---
title: "Phase 1: Retrieval of Religious Tourism Destination Corpus"
subtitle: "Per-destination matching with context gate, religious register tagging"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(duckdb)
library(DBI)
library(data.table)
library(stringi)
library(ggplot2)
library(knitr)
library(patchwork)

dirs <- c("data-raw", "data", "output", "output/tables", "output/figures",
          "output/reports", "dictionaries", "R")
for (d in dirs) if (!dir.exists(d)) dir.create(d, recursive = TRUE)
```

# Design Notes

The retrieval pipeline extracts a destination-level working corpus from the
Determ database of Croatian digital media (2021 to May 2024). Key design
choices are documented in the companion file `01_retrieval_logic.md`. The
summary of decisions implemented here is as follows.

1. Matching uses TITLE and MENTION_SNIPPET only, not FULL_TEXT. FULL_TEXT
   is retrieved for downstream text modelling but is not a filter target.
2. Each destination is queried separately. Results are assembled into a
   long-format table where each row is one article-destination pair.
3. Polysemic destinations (Sinj, Nin, Solin, Krasno, Ludbreg) are subject
   to a context gate. Matching requires both a destination pattern match
   and a context trigger match in the same article (TITLE or snippet).
4. Unambiguous destinations (Marija Bistrica, Trsat, Aljmaš, Voćin, Vepric)
   are matched without a context gate.
5. Religious, tourism, and scandal registers are tagged on FULL_TEXT after
   retrieval, as binary indicators for downstream framing analysis.
6. All cases are applied in full for case-sensitive Croatian morphology.

# 0. Configuration

```{r}
#| label: config

config <- list(
  duckdb_path       = "C:/Users/lsikic/Luka C/DetermDB/determDB.duckdb",
  db_memory_limit   = "48GB",
  seed              = 2026L,
  analysis_start    = as.Date("2021-01-01"),
  analysis_end      = as.Date("2024-05-31"),
  language_filter   = c("hr", "bs"),
  min_articles_stm  = 300L,
  min_articles_desc = 50L,
  minimum_months    = 41L
)
saveRDS(config, "data/config_destinations.rds")
cat("Config saved. Analysis window:", format(config$analysis_start),
    "to", format(config$analysis_end), "\n")
```

# 1. Destination Dictionary

Each destination has a name, a type classification, a vector of ILIKE
patterns covering Croatian case forms and common shrine naming variants,
a flag indicating whether the context gate applies, and a vector of
context triggers (used only when the gate applies).

```{r}
#| label: destination-dictionary

destinations <- list(

  marija_bistrica = list(
    type = "nacionalno_marijansko",
    context_required = FALSE,
    patterns = c(
      "marija bistric",          # Marija Bistrica, Mariji Bistrici etc
      "mariju bistricu",
      "marije bistrice",
      "marijom bistricom",
      "bistričk",                # bistrička Gospa, bistričkog svetišta
      "majka božja bistrička",
      "majke božje bistričke",
      "majku božju bistričku"
    ),
    context_triggers = character(0)
  ),

  trsat = list(
    type = "nacionalno_marijansko",
    context_required = FALSE,
    patterns = c(
      "trsat",                   # Trsat, Trsata, Trsatu, Trsatom
      "trsatsk",                 # trsatski, trsatska, trsatsko
      "gospa trsatsk",
      "gospu trsatsk",
      "gospe trsatsk",
      "gospom trsatsk",
      "majka božja trsatsk"
    ),
    context_triggers = character(0)
  ),

  sinj = list(
    type = "nacionalno_marijansko",
    context_required = TRUE,
    patterns = c(
      "sinj",                    # Sinj, Sinja, Sinju, Sinjem
      "sinjsk",                  # sinjska, sinjsko, sinjski, sinjskoga
      "alka",                    # Alka, Alke, Alku, Alkom
      "alkar"                    # alkari, alkara
    ),
    context_triggers = c(
      "gospa", "svetiš", "hodočast", "marij", "sveta",
      "čudotvor", "kip", "slik", "crkv", "fratri", "franjev",
      "alka", "alkar", "tradicij", "procesij", "cetinsk"
    )
  ),

  aljmas = list(
    type = "lokalno_marijansko",
    context_required = FALSE,
    patterns = c(
      "aljmaš",
      "aljmaša",
      "aljmašu",
      "aljmašem",
      "aljmašk",                 # aljmaška Gospa
      "gospa aljmašk",
      "gospe aljmašk",
      "gospu aljmašk",
      "utočišt",                 # Gospa Utočišta grešnika (aljmaški titular)
      "majka božja aljmašk"
    ),
    context_triggers = character(0)
  ),

  ludbreg = list(
    type = "lokalno_marijansko",
    context_required = TRUE,
    patterns = c(
      "ludbreg",                 # Ludbreg, Ludbrega, Ludbregu
      "ludbrešk",                # ludbreški, ludbreška, ludbreškom
      "ludbrešku"
    ),
    context_triggers = c(
      "sveta krv", "svete krvi", "svetu krv", "svetom krvi",
      "euharistij", "čudo", "čuda", "čudesn",
      "svetiš", "hodočast", "marij", "procesij",
      "relikvij", "marijansk"
    )
  ),

  krasno = list(
    type = "lokalno_marijansko",
    context_required = TRUE,
    patterns = c(
      "krasn",                   # Krasno, Krasna, Krasnu, Krasnom
      "krasnarsk"                # rare variant
    ),
    context_triggers = c(
      "gospa od krasn", "gospe od krasn", "gospu od krasn", "gospom od krasn",
      "senjsk", "marijansk", "svetiš", "hodočast",
      "hrvatsko svetišt", "velebit"
    )
  ),

  vocin = list(
    type = "lokalno_marijansko",
    context_required = FALSE,
    patterns = c(
      "voćin",                   # Voćin, Voćina, Voćinu, Voćinom
      "voćinsk",                 # voćinska Gospa
      "gospa voćinsk",
      "gospe voćinsk",
      "gospu voćinsk",
      "gospom voćinsk",
      "majka božja voćinsk"
    ),
    context_triggers = character(0)
  ),

  vepric = list(
    type = "lokalno_marijansko",
    context_required = FALSE,
    patterns = c(
      "vepric",                  # Vepric, Veprica, Vepricu
      "veprič",                  # veprički
      "gospa od lurda",          # Vepric is known as Croatian Lourdes
      "gospe od lurda",
      "gospu od lurda",
      "hrvatski lurd"
    ),
    context_triggers = character(0)
  ),

  nin = list(
    type = "nemarijansko",
    context_required = TRUE,
    patterns = c(
      "nin",                     # Nin, Nina, Ninu, Ninom
      "ninsk",                   # ninska, ninski, ninsko
      "ninskim"
    ),
    context_triggers = c(
      "tavelić", "nikola tavel",
      "bazilik", "katedral",
      "starohrvatsk", "hrvatsk", "branimir", "tomislav",
      "svet", "crkv", "biskup",
      "gospa", "hodočast",
      "zavjet", "krunidb"
    )
  ),

  solin = list(
    type = "nemarijansko",
    context_required = TRUE,
    patterns = c(
      "solin",                   # Solin, Solina, Solinu, Solinom
      "solinsk"                  # solinska, solinski
    ),
    context_triggers = c(
      "gospa od otoka", "gospe od otoka", "gospu od otoka", "gospom od otoka",
      "starohrvatsk", "salonitansk", "salon",
      "arheolog", "iskopin", "ruševin",
      "crkv", "bazilik", "kralj zvonim",
      "hodočast", "marijansk", "svetiš",
      "šuplj crkv", "gradin"
    )
  )
)

cat("Destinations configured:", length(destinations), "\n")
cat("  Type breakdown:\n")
type_summary <- table(sapply(destinations, function(d) d$type))
print(type_summary)

cat("\n  Context gate required:\n")
gate_summary <- data.table(
  destination = names(destinations),
  type = sapply(destinations, function(d) d$type),
  context_gate = sapply(destinations, function(d) d$context_required),
  n_patterns = sapply(destinations, function(d) length(d$patterns)),
  n_triggers = sapply(destinations, function(d) length(d$context_triggers))
)
kable(gate_summary, caption = "Destination dictionary structure")
```

# 2. Register Dictionaries

Register dictionaries are applied after retrieval, tagging each article
on FULL_TEXT. Reduced Catholic root set derived from DigiKat project.

```{r}
#| label: register-dictionaries

registers <- list(

  religious = c(
    "svetiš", "hodočast", "hodočasn",
    "biskup", "nadbiskup", "kardinal", "svećenik", "fratar", "franjev",
    "papa ", "papinsk", "vatikan",
    "misa ", "misu ", "mise ", "misom",
    "liturgij", "euharistij", "pričes",
    "gospa", "majka božj", "isus", "krist",
    "molitv", "blagoslov", "sakram",
    "katolič", "kršćan", "vjernic", "vjera ",
    "crkv", "katedral", "kapel", "bazilik",
    "blagdan", "korizm", "advent", "uskrs",
    "procesij", "zavjet", "čudo", "čuda", "čudotvor",
    "marijansk", "duhovn"
  ),

  tourism = c(
    "turizam", "turist", "turistič",
    "destinacij", "posjetitelj", "posjet",
    "dolasci", "dolazak",
    "noćenj", "smještaj",
    "hotel", "apartm",
    "ljeto", "ljetn", "sezon",
    "hrvatska turistič", "turistička zajednic",
    "putnic", "putovanj",
    "izletnic", "izlet"
  ),

  scandal = c(
    "nesreć", "poginuo", "poginul", "stradal",
    "skandal", "afer",
    "istraga", "optužn", "sud",
    "korupcij", "pronevjer", "zloporab",
    "sukob", "kriminal",
    "pljačk", "krađ",
    "prevar", "šok"
  )
)

cat("Register dictionaries:\n")
for (nm in names(registers)) {
  cat("  ", nm, ":", length(registers[[nm]]), "root patterns\n")
}
```

# 3. Connect to Database

```{r}
#| label: connect

con <- dbConnect(duckdb::duckdb(), dbdir = config$duckdb_path, read_only = TRUE)
dbExecute(con, paste0("SET memory_limit='", config$db_memory_limit, "';"))
main_tbl <- "media_data"
cat("Connected. Table:", main_tbl, "\n")

# verify columns we need exist
cols <- dbGetQuery(con, paste0("SELECT * FROM \"", main_tbl, "\" LIMIT 0"))
needed <- c("DATE", "TITLE", "FROM", "SOURCE_TYPE", "REACH",
            "AUTO_SENTIMENT", "INTERACTIONS", "MENTION_SNIPPET",
            "FULL_TEXT", "URL", "LANGUAGES")
missing <- setdiff(needed, colnames(cols))
if (length(missing) > 0) {
  stop("Missing columns in database: ", paste(missing, collapse = ", "))
}
cat("All required columns present.\n")
```

# 4. Query Builders

```{r}
#| label: query-builders

# build OR block of ILIKE clauses over TITLE and MENTION_SNIPPET
build_ilike_block <- function(patterns) {
  title_clauses <- paste0("\"TITLE\" ILIKE '%", patterns, "%'")
  snippet_clauses <- paste0("\"MENTION_SNIPPET\" ILIKE '%", patterns, "%'")
  paste(c(title_clauses, snippet_clauses), collapse = "\n    OR ")
}

# build full WHERE clause for one destination, applying context gate if needed
build_destination_where <- function(dest_entry) {
  dest_block <- build_ilike_block(dest_entry$patterns)

  if (dest_entry$context_required) {
    ctx_block <- build_ilike_block(dest_entry$context_triggers)
    where_core <- paste0(
      "(\n    ", dest_block, "\n  )\n  AND (\n    ", ctx_block, "\n  )"
    )
  } else {
    where_core <- paste0("(\n    ", dest_block, "\n  )")
  }

  date_clause <- paste0(
    "CAST(\"DATE\" AS DATE) >= '", config$analysis_start, "'\n  AND ",
    "CAST(\"DATE\" AS DATE) <= '", config$analysis_end, "'"
  )

  # language filter (hr or bs)
  lang_clauses <- paste0("\"LANGUAGES\" ILIKE '%", config$language_filter, "%'")
  lang_block <- paste(lang_clauses, collapse = " OR ")

  paste0(where_core, "\nAND (", date_clause, ")\nAND (", lang_block, ")")
}

# preview example query
example_where <- build_destination_where(destinations$sinj)
cat("Example WHERE clause for Sinj (context gate active):\n")
cat(substr(example_where, 1, 800), "...\n")
```

# 5. Per-Destination Retrieval

Loop through destinations, execute one query per destination, collect
results. Each result is tagged with its destination label.

```{r}
#| label: retrieve-loop

select_cols <- "\"DATE\", \"TITLE\", \"FROM\", \"SOURCE_TYPE\",
\"REACH\", \"AUTO_SENTIMENT\", \"INTERACTIONS\",
\"LIKE_COUNT\", \"COMMENT_COUNT\", \"SHARE_COUNT\",
\"MENTION_SNIPPET\", \"FULL_TEXT\", \"URL\", \"LANGUAGES\""

dest_results <- list()
query_log <- data.table()

for (dname in names(destinations)) {
  dentry <- destinations[[dname]]
  where_sql <- build_destination_where(dentry)

  q <- paste0("SELECT ", select_cols, "\nFROM \"", main_tbl,
              "\"\nWHERE ", where_sql)

  t0 <- Sys.time()
  res <- as.data.table(dbGetQuery(con, q))
  t1 <- Sys.time()

  res[, destination := dname]
  res[, dest_type := dentry$type]
  res[, context_gate_applied := dentry$context_required]

  dest_results[[dname]] <- res

  log_row <- data.table(
    destination = dname,
    type = dentry$type,
    context_gate = dentry$context_required,
    n_patterns = length(dentry$patterns),
    n_triggers = length(dentry$context_triggers),
    articles_matched = nrow(res),
    query_seconds = round(as.numeric(difftime(t1, t0, units = "secs")), 1)
  )
  query_log <- rbind(query_log, log_row)

  cat(sprintf("  %-18s | %6s articles | %5.1fs | gate=%s\n",
              dname, format(nrow(res), big.mark = ","),
              log_row$query_seconds,
              ifelse(dentry$context_required, "Y", "N")))
}

cat("\nRetrieval complete. Summary:\n")
kable(query_log, format.args = list(big.mark = ","),
      caption = "Per-destination retrieval summary")
```

# 6. Assemble Long Format

```{r}
#| label: assemble-long

long <- rbindlist(dest_results, use.names = TRUE, fill = TRUE)
long[, date := as.Date(DATE)]
long[, yearmonth := as.Date(format(date, "%Y-%m-01"))]
long[, year := year(date)]
long[, week := as.Date(cut(date, "week"))]
long[, dow := weekdays(date)]
long[, pair_id := .I]

# a unique article hash for deduplication across destinations
# uses URL if available, otherwise FROM+DATE+TITLE hash
long[, article_key := ifelse(is.na(URL) | URL == "",
                              paste(FROM, DATE, substr(TITLE, 1, 50), sep = "|"),
                              URL)]
long[, article_id := .GRP, by = article_key]

cat("Long-format table assembled.\n")
cat("  Rows (article-destination pairs):", format(nrow(long), big.mark = ","), "\n")
cat("  Unique articles:", format(uniqueN(long$article_id), big.mark = ","), "\n")
cat("  Articles mentioning multiple destinations:",
    format(sum(long[, .N, by = article_id][, N > 1]), big.mark = ","), "\n")

# multi-destination article breakdown
n_dest_per_article <- long[, .(n_dest = uniqueN(destination)), by = article_id]
multi_dest_tab <- n_dest_per_article[, .N, by = n_dest][order(n_dest)]
kable(multi_dest_tab,
      col.names = c("Destinations per article", "Articles"),
      format.args = list(big.mark = ","),
      caption = "Co-mention distribution")
```

# 7. Register Tagging on FULL_TEXT

```{r}
#| label: register-tagging

# build one big regex per register for efficient matching
build_register_pattern <- function(terms) {
  # escape any regex metacharacters in terms (safety)
  escaped <- stri_replace_all_regex(terms, "([.\\\\+*?\\[\\]{}^$()|])", "\\\\$1")
  paste0("(?i)(", paste(escaped, collapse = "|"), ")")
}

pat_religious <- build_register_pattern(registers$religious)
pat_tourism   <- build_register_pattern(registers$tourism)
pat_scandal   <- build_register_pattern(registers$scandal)

cat("Tagging registers on FULL_TEXT (this may take a minute)...\n")

t0 <- Sys.time()
long[, has_religious_register := stri_detect_regex(FULL_TEXT, pat_religious) |
                                   stri_detect_regex(TITLE, pat_religious)]
long[, has_tourism_register   := stri_detect_regex(FULL_TEXT, pat_tourism) |
                                   stri_detect_regex(TITLE, pat_tourism)]
long[, has_scandal_register   := stri_detect_regex(FULL_TEXT, pat_scandal) |
                                   stri_detect_regex(TITLE, pat_scandal)]
t1 <- Sys.time()

cat("Register tagging done in",
    round(as.numeric(difftime(t1, t0, units = "secs")), 1), "seconds.\n\n")

# handle NAs (FULL_TEXT missing)
for (col in c("has_religious_register", "has_tourism_register",
              "has_scandal_register")) {
  long[is.na(get(col)), (col) := FALSE]
}

register_summary <- long[, .(
  n_pairs = .N,
  pct_religious = round(mean(has_religious_register) * 100, 1),
  pct_tourism   = round(mean(has_tourism_register) * 100, 1),
  pct_scandal   = round(mean(has_scandal_register) * 100, 1)
), by = .(destination, dest_type)][order(-n_pairs)]

kable(register_summary, format.args = list(big.mark = ","),
      caption = "Register prevalence by destination")
```

# 8. Quality Profile

```{r}
#| label: quality-profile

cat("=== Overall corpus quality ===\n\n")
cat("Total article-destination pairs:", format(nrow(long), big.mark = ","), "\n")
cat("Unique articles:", format(uniqueN(long$article_id), big.mark = ","), "\n")
cat("Unique sources:", format(uniqueN(long$FROM), big.mark = ","), "\n")
cat("Date range:", format(min(long$date)), "to", format(max(long$date)), "\n")
cat("Months covered:", uniqueN(long$yearmonth),
    "of", config$minimum_months, "expected\n\n")

# platform distribution per destination
platform_dest <- long[, .N, by = .(destination, SOURCE_TYPE)]
platform_dest_wide <- dcast(platform_dest, destination ~ SOURCE_TYPE,
                             value.var = "N", fill = 0L)
kable(platform_dest_wide, format.args = list(big.mark = ","),
      caption = "Article-destination pairs by platform")

# monthly coverage
monthly_cov <- long[, .(n_pairs = .N, n_articles = uniqueN(article_id)),
                     by = yearmonth][order(yearmonth)]
cat("Thinnest month:", format(monthly_cov[which.min(n_pairs)]$yearmonth),
    "with", min(monthly_cov$n_pairs), "pairs\n")
cat("Thickest month:", format(monthly_cov[which.max(n_pairs)]$yearmonth),
    "with", max(monthly_cov$n_pairs), "pairs\n")
cat("Mean pairs per month:", round(mean(monthly_cov$n_pairs)), "\n")

# sentiment distribution
cat("\nAuto sentiment distribution (on pairs):\n")
sent_tab <- long[, .N, by = AUTO_SENTIMENT][order(-N)]
sent_tab[, pct := round(N / sum(N) * 100, 1)]
print(sent_tab)
```

```{r}
#| label: monthly-volume-plot

p_vol <- ggplot(monthly_cov, aes(yearmonth, n_pairs)) +
  geom_col(fill = "#2E4057", alpha = 0.85) +
  geom_hline(yintercept = 50, linetype = "dotted", colour = "orange") +
  labs(title = "Monthly article-destination pairs",
       subtitle = "Dotted line marks 50-pair threshold",
       x = NULL, y = "Pairs") +
  theme_minimal()
ggsave("output/figures/01_monthly_volume_destinations.png",
       p_vol, width = 10, height = 5)
p_vol
```

```{r}
#| label: per-dest-timeseries

monthly_dest <- long[, .(n = .N),
                      by = .(destination, dest_type, yearmonth)]

p_ts <- ggplot(monthly_dest, aes(yearmonth, n, colour = dest_type)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~ destination, scales = "free_y", ncol = 2) +
  labs(title = "Monthly volume by destination",
       x = NULL, y = "Pairs", colour = "Type") +
  theme_minimal() +
  theme(legend.position = "bottom")
ggsave("output/figures/01_per_destination_timeseries.png",
       p_ts, width = 12, height = 10)
p_ts
```

# 9. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 1\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. all destinations returned non-empty
dest_nonempty <- query_log$articles_matched > 0
checks$all_destinations_present <- all(dest_nonempty)
cat("1. All destinations returned articles:",
    ifelse(checks$all_destinations_present, "PASS", "FAIL"), "\n")
if (!checks$all_destinations_present) {
  empty <- query_log[articles_matched == 0]$destination
  cat("   Empty:", paste(empty, collapse = ", "), "\n")
}

# 2. every destination meets descriptive threshold
above_desc <- query_log$articles_matched >= config$min_articles_desc
checks$all_above_descriptive <- all(above_desc)
cat("2. All destinations >= ", config$min_articles_desc, " articles: ",
    ifelse(checks$all_above_descriptive, "PASS", "CAUTION"), "\n", sep = "")
if (!checks$all_above_descriptive) {
  thin <- query_log[articles_matched < config$min_articles_desc]$destination
  cat("   Below threshold:", paste(thin, collapse = ", "), "\n")
}

# 3. count destinations meeting STM threshold
n_above_stm <- sum(query_log$articles_matched >= config$min_articles_stm)
checks$stm_ready <- n_above_stm >= 6
cat("3. At least 6 destinations >= ", config$min_articles_stm,
    " articles (STM-ready): ",
    ifelse(checks$stm_ready, "PASS", "CAUTION"),
    " (", n_above_stm, " of ", length(destinations), ")\n", sep = "")

# 4. month coverage
checks$months_full <- uniqueN(long$yearmonth) >= config$minimum_months
cat("4. All ", config$minimum_months, " months covered: ",
    ifelse(checks$months_full, "PASS", "FAIL"),
    " (", uniqueN(long$yearmonth), ")\n", sep = "")

# 5. religious register prevalence in plausible range
overall_religious <- mean(long$has_religious_register)
checks$religious_ok <- overall_religious > 0.20 & overall_religious < 0.95
cat("5. Religious register between 20-95% overall: ",
    ifelse(checks$religious_ok, "PASS", "NOTE"),
    " (", round(overall_religious * 100, 1), "%)\n", sep = "")

# 6. platform diversity (at least 3 platforms per major destination)
major_dests <- query_log[articles_matched >= config$min_articles_stm]$destination
plat_per_major <- long[destination %in% major_dests,
                         .(n_plat = uniqueN(SOURCE_TYPE)), by = destination]
checks$platform_div <- all(plat_per_major$n_plat >= 3)
cat("6. Major destinations cover 3+ platforms: ",
    ifelse(checks$platform_div, "PASS", "NOTE"), "\n", sep = "")

# 7. language filter working (hr dominant)
lang_sample <- stri_count_regex(long$LANGUAGES, "hr") > 0
checks$language_ok <- mean(lang_sample, na.rm = TRUE) > 0.90
cat("7. Croatian language > 90% of matches: ",
    ifelse(checks$language_ok, "PASS", "NOTE"),
    " (", round(mean(lang_sample, na.rm = TRUE) * 100, 1), "%)\n", sep = "")

cat("\n")
n_pass <- sum(unlist(checks))
cat("Passed:", n_pass, "of", length(checks), "\n")
if (all(unlist(checks))) {
  cat("ALL CHECKS PASSED. Proceed to Phase 2.\n")
} else {
  failed <- names(checks)[!unlist(checks)]
  cat("Attention needed:", paste(failed, collapse = ", "), "\n")
}
cat(strrep("=", 60), "\n")
```

# 10. Disconnect

```{r}
#| label: disconnect

dbDisconnect(con, shutdown = TRUE)
cat("Database connection closed.\n")
```

# 11. Save Outputs

```{r}
#| label: save-outputs

# main long-format corpus
saveRDS(long, "data-raw/destinations_long.rds")

# per-destination list for inspection
saveRDS(dest_results, "data-raw/destinations_by_destination.rds")

# query log
saveRDS(query_log, "data/destinations_query_log.rds")
fwrite(query_log, "output/tables/destinations_query_log.csv")

# register summary
saveRDS(register_summary, "data/destinations_register_summary.rds")
fwrite(register_summary, "output/tables/destinations_register_summary.csv")

# file sizes
cat("Output files written:\n")
fs <- c(
  "data-raw/destinations_long.rds",
  "data-raw/destinations_by_destination.rds",
  "data/destinations_query_log.rds",
  "data/destinations_register_summary.rds"
)
for (f in fs) {
  if (file.exists(f)) {
    cat("  ", f, "|", round(file.size(f) / 1e6, 1), "MB\n")
  }
}
```

# 12. Metadata Snapshot

```{r}
#| label: metadata

retrieval_meta <- list(
  timestamp       = Sys.time(),
  phase           = "1_destination_retrieval",
  r_version       = R.version.string,
  duckdb_version  = packageVersion("duckdb"),
  config          = config,
  destinations    = destinations,
  registers       = registers,
  query_log       = query_log,
  register_summary = register_summary,
  gate_results    = checks,
  n_pairs         = nrow(long),
  n_articles      = uniqueN(long$article_id),
  months_covered  = uniqueN(long$yearmonth)
)

saveRDS(retrieval_meta, "data/destinations_metadata.rds")
cat("Metadata saved.\n\n")
cat("Phase 1 complete. Primary working corpus: destinations_long.rds\n")
cat("Next step: Phase 2 descriptive analysis.\n")
```
