---
title: "Phase 1b: Diagnostic Validation of Sensitive Destinations"
subtitle: "Contamination check for Sinj, Nin, Solin before dictionary refinement"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(stringi)
library(ggplot2)
library(knitr)
library(DT)

for (d in c("output/tables", "output/figures", "output/reports")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Purpose

Phase 1 retrieval produced an asymmetric corpus dominated by Sinj (119,330)
and Nin (95,507). Religious register prevalence for these two destinations
is far below the corpus mean, suggesting that the context gate admitted
articles that are not actually about the religious tourism destinations.
This script performs three diagnostics.

1. Random snippet sampling from Sinj, Nin, and Solin to eyeball actual
   content.
2. Frequency analysis of surrounding tokens to identify dominant non-target
   contexts (person names, sports, music, other polysemic uses).
3. Cross tabulation of register indicators against destination matches to
   quantify how many matches lack any religious or tourism anchor.

Outputs inform a binary decision between two refinement strategies.
Strategy A tightens destination patterns and context triggers. Strategy B
adds a register forced gate (destination match AND religious/tourism
register match) and accepts that descriptive coverage narrows to the
religious tourism discourse itself.

```{r}
#| label: load-corpus

long <- readRDS("data-raw/destinations_long.rds")
cat("Loaded long format corpus.\n")
cat("  Rows:", format(nrow(long), big.mark = ","), "\n")
cat("  Columns:", ncol(long), "\n\n")
```

# 1. Random Snippet Sampling

```{r}
#| label: random-samples

set.seed(2026)

sensitive_dests <- c("sinj", "nin", "solin")
n_per <- 50

samples <- list()
for (d in sensitive_dests) {
  subset_d <- long[destination == d]
  if (nrow(subset_d) > n_per) {
    samples[[d]] <- subset_d[sample(.N, n_per)]
  } else {
    samples[[d]] <- subset_d
  }
}

for (d in sensitive_dests) {
  cat("=== ", toupper(d), " (", format(nrow(samples[[d]]), big.mark = ","),
      " sampled) ===\n", sep = "")
  cat("(", format(nrow(long[destination == d]), big.mark = ","),
      " total in corpus)\n\n", sep = "")

  display <- samples[[d]][, .(
    date = format(date),
    source = substr(FROM, 1, 25),
    title = substr(TITLE, 1, 80),
    snippet = substr(MENTION_SNIPPET, 1, 200),
    religious = has_religious_register,
    tourism = has_tourism_register,
    scandal = has_scandal_register
  )]

  fwrite(display, paste0("output/tables/01b_sample_", d, ".csv"))

  # print first 15 rows for quick visual check
  for (i in 1:min(15, nrow(display))) {
    cat("[", i, "] ", display$date[i], " | ", display$source[i], "\n", sep = "")
    cat("    TITLE:  ", display$title[i], "\n", sep = "")
    cat("    SNIP:   ", display$snippet[i], "\n", sep = "")
    cat("    REG:    rel=", display$religious[i],
        " tour=", display$tourism[i],
        " scan=", display$scandal[i], "\n\n", sep = "")
  }
  cat("\n--- end of ", d, " sample. Full 50 in output/tables/01b_sample_", d, ".csv ---\n\n\n", sep = "")
}
```

# 2. Register Cross Tabulation

```{r}
#| label: register-crosstab

# articles with NO register hit at all
register_presence <- long[, .(
  n_pairs = .N,
  no_register = sum(!has_religious_register & !has_tourism_register & !has_scandal_register),
  religious_only = sum(has_religious_register & !has_tourism_register & !has_scandal_register),
  tourism_only = sum(!has_religious_register & has_tourism_register & !has_scandal_register),
  scandal_only = sum(!has_religious_register & !has_tourism_register & has_scandal_register),
  multi = sum(has_religious_register + has_tourism_register + has_scandal_register > 1)
), by = destination][order(-n_pairs)]

register_presence[, pct_no_register := round(no_register / n_pairs * 100, 1)]
register_presence[, pct_religious_only := round(religious_only / n_pairs * 100, 1)]
register_presence[, pct_tourism_only := round(tourism_only / n_pairs * 100, 1)]
register_presence[, pct_scandal_only := round(scandal_only / n_pairs * 100, 1)]
register_presence[, pct_multi := round(multi / n_pairs * 100, 1)]

kable(register_presence[, .(destination, n_pairs,
                             pct_no_register, pct_religious_only,
                             pct_tourism_only, pct_scandal_only, pct_multi)],
      format.args = list(big.mark = ","),
      caption = "Register distribution per destination (% of pairs)")
```

```{r}
#| label: no-register-analysis

# of all Sinj and Nin pairs, how many have NO register at all?
no_reg_sensitive <- long[destination %in% sensitive_dests &
                           !has_religious_register &
                           !has_tourism_register &
                           !has_scandal_register]

cat("Sensitive destinations pairs with NO register match:\n")
cat("  Total:", format(nrow(no_reg_sensitive), big.mark = ","), "\n")
cat("  By destination:\n")
print(no_reg_sensitive[, .N, by = destination])
```

# 3. Context Token Frequency Around Destination Matches

For the large Sinj and Nin corpora, extract the most frequent tokens
appearing in MENTION_SNIPPET. This reveals what the articles are
actually about.

```{r}
#| label: token-frequency

library(quanteda)

extract_top_tokens <- function(dest_name, n_top = 40) {
  subset_d <- long[destination == dest_name]
  snippets <- subset_d$MENTION_SNIPPET
  snippets <- snippets[!is.na(snippets)]

  # quick tokenisation
  toks <- tokens(snippets, what = "word",
                 remove_punct = TRUE, remove_numbers = TRUE,
                 remove_symbols = TRUE, remove_url = TRUE) |>
    tokens_tolower() |>
    tokens_select(min_nchar = 3)

  # remove trivial Croatian stopwords
  common_stops <- c("koji", "koja", "koje", "kako", "kada", "gdje", "tako",
                    "samo", "više", "nego", "jedan", "dva", "tri", "sve",
                    "bila", "bilo", "bile", "biti", "bila", "onda", "ovaj",
                    "ova", "ovo", "njezin", "njegov", "sebe", "sobom",
                    "svih", "svog", "svoje", "svim", "svu", "tome", "tog",
                    "prema", "među", "takvim", "njima", "pri", "radi")
  toks <- tokens_remove(toks, common_stops)

  dfm_d <- dfm(toks)
  top <- topfeatures(dfm_d, n = n_top)
  data.table(token = names(top), freq = as.integer(top))
}

for (d in sensitive_dests) {
  cat("=== Top tokens in", toupper(d), "snippets ===\n\n")
  top_tokens <- extract_top_tokens(d, 40)
  print(top_tokens)
  fwrite(top_tokens, paste0("output/tables/01b_top_tokens_", d, ".csv"))
  cat("\n")
}
```

# 4. Known Contamination Patterns

Specific regex checks for known non-target uses. These tell us directly
how large the contamination is.

```{r}
#| label: known-contamination

contamination_checks <- function(dest_name) {
  subset_d <- long[destination == dest_name]
  snip <- paste(subset_d$MENTION_SNIPPET, subset_d$TITLE, sep = " | ")

  if (dest_name == "sinj") {
    patterns <- c(
      "Sinjković"           = "sinjković",
      "Junak Sinj (sport)"  = "junak sinj|nk sinj|rk sinj",
      "alkarsko (non shrine)" = "alkarsk[aeiou]",
      "prometna nesreća u Sinju" = "nesreć.{0,50}(u|kod|blizu).{0,10}sinj",
      "bez bilo kojeg religijskog trigera" = "^(?!.*(gospa|svet|marij|crkv|fratr|hodočast|franjev)).*$"
    )
  } else if (dest_name == "nin") {
    patterns <- c(
      "Nina (ime)"          = "\\bnin[aey]\\b",
      "Nino (ime)"          = "\\bnin[oa]\\s+[A-ZČĆŽŠĐ]",
      "nine (eng num)"      = "\\bnine\\b",
      "training/mining"     = "training|mining|dining|burning",
      "bez religijskog trigera" = "^(?!.*(tavel|bazilik|starohrvat|zavjet|krunidb|svetiš)).*$"
    )
  } else if (dest_name == "solin") {
    patterns <- c(
      "Solinjanin (prezime)" = "solinjanin",
      "HNK Solin"           = "hnk solin|nk solin",
      "bez religijskog trigera" = "^(?!.*(gospa od otoka|starohrvat|svetiš|bazilik)).*$"
    )
  }

  results <- data.table(
    pattern_name = names(patterns),
    n_matches = sapply(patterns, function(p) {
      sum(stri_detect_regex(snip, p, opts_regex = list(case_insensitive = TRUE)), na.rm = TRUE)
    })
  )
  results[, total := nrow(subset_d)]
  results[, pct := round(n_matches / total * 100, 1)]
  results
}

for (d in sensitive_dests) {
  cat("=== Contamination patterns in", toupper(d), "===\n")
  cont <- contamination_checks(d)
  print(kable(cont, caption = paste("Contamination check:", d)))
  cat("\n")
}
```

# 5. Control Comparison

Compare register prevalence in sensitive destinations against unambiguous
destinations (Marija Bistrica, Trsat, Aljmaš, Voćin, Vepric) to establish
a baseline for what a clean corpus looks like.

```{r}
#| label: clean-vs-sensitive

clean_dests <- c("marija_bistrica", "trsat", "aljmas", "vocin", "vepric")

comparison <- long[, .(
  n_pairs = .N,
  pct_religious = round(mean(has_religious_register) * 100, 1),
  pct_tourism = round(mean(has_tourism_register) * 100, 1),
  pct_any_target_register = round(mean(has_religious_register | has_tourism_register) * 100, 1)
), by = .(destination, dest_type)]

comparison[, group := fifelse(destination %in% sensitive_dests, "sensitive",
                              fifelse(destination %in% clean_dests, "clean", "other"))]
comparison <- comparison[order(group, -n_pairs)]

kable(comparison,
      format.args = list(big.mark = ","),
      caption = "Clean vs sensitive destination comparison")

cat("\nGROUP MEANS (weighted by pair count):\n")
group_means <- comparison[, .(
  total_pairs = sum(n_pairs),
  weighted_religious = round(sum(pct_religious * n_pairs) / sum(n_pairs), 1),
  weighted_tourism = round(sum(pct_tourism * n_pairs) / sum(n_pairs), 1),
  weighted_any_target = round(sum(pct_any_target_register * n_pairs) / sum(n_pairs), 1)
), by = group]
print(group_means)
```

# 6. Recommendation Generator

Based on the diagnostic numbers, produce a recommendation for each
sensitive destination.

```{r}
#| label: recommendation

recommendations <- data.table(
  destination = sensitive_dests
)

for (d in sensitive_dests) {
  pairs <- long[destination == d, .N]
  pct_rel <- long[destination == d, mean(has_religious_register) * 100]
  pct_tour <- long[destination == d, mean(has_tourism_register) * 100]
  pct_any <- long[destination == d, mean(has_religious_register | has_tourism_register) * 100]

  recommendations[destination == d, `:=`(
    pairs = pairs,
    pct_religious = round(pct_rel, 1),
    pct_tourism = round(pct_tour, 1),
    pct_any_target = round(pct_any, 1),
    verdict = fcase(
      pct_any < 30, "HEAVY CONTAMINATION - Strategy B (register forced gate)",
      pct_any < 50, "MODERATE CONTAMINATION - Strategy A or B",
      pct_any < 70, "MILD CONTAMINATION - Strategy A (tighten triggers)",
      default = "ACCEPTABLE - keep as is"
    ),
    projected_after_B = round(pairs * pct_any / 100)
  )]
}

kable(recommendations, format.args = list(big.mark = ","),
      caption = "Per-destination recommendation")
```

# 7. Summary Decision Brief

```{r}
#| label: summary

cat(strrep("=", 60), "\n")
cat("DIAGNOSTIC SUMMARY\n")
cat(strrep("=", 60), "\n\n")

total_sensitive <- sum(recommendations$pairs)
projected_B <- sum(recommendations$projected_after_B)

cat("Current sensitive corpus size:", format(total_sensitive, big.mark = ","),
    "pairs\n")
cat("Projected after Strategy B filter:",
    format(projected_B, big.mark = ","),
    " pairs (",
    round(projected_B / total_sensitive * 100, 1), "%)\n\n", sep = "")

cat("Clean destinations total:",
    format(long[destination %in% clean_dests, .N], big.mark = ","), "\n")
cat("Ludbreg + Krasno total:",
    format(long[destination %in% c("ludbreg", "krasno"), .N], big.mark = ","),
    "\n\n")

cat("IMPLICATIONS FOR FINAL CORPUS SIZE:\n\n")
cat("Strategy A (tighten patterns + triggers):\n")
cat("  Unpredictable. Depends on how many genuine articles are kept.\n")
cat("  Likely 15-30K total for sensitive destinations.\n\n")

cat("Strategy B (register forced gate for sensitive only):\n")
cat("  Projected ", format(projected_B, big.mark = ","),
    " sensitive pairs\n")
cat("  Clean destinations unaffected\n")
cat("  Total corpus ~ ",
    format(long[!destination %in% sensitive_dests, .N] + projected_B,
           big.mark = ","), " pairs\n\n", sep = "")

cat("Strategy C (register forced gate globally):\n")
total_C <- long[has_religious_register | has_tourism_register, .N]
cat("  Applies register gate to ALL destinations\n")
cat("  Total corpus size: ", format(total_C, big.mark = ","),
    " pairs (",
    round(total_C / nrow(long) * 100, 1), "% of current)\n", sep = "")
cat("  Cleanest but loses some legitimate no register mentions\n")
cat(strrep("=", 60), "\n")
```

# 8. Save Diagnostics

```{r}
#| label: save

diagnostic_out <- list(
  timestamp = Sys.time(),
  samples = samples,
  register_presence = register_presence,
  comparison = comparison,
  group_means = group_means,
  recommendations = recommendations,
  projected_strategy_B = projected_B,
  total_strategy_C = total_C
)
saveRDS(diagnostic_out, "data/destinations_diagnostic.rds")
cat("Diagnostic saved to data/destinations_diagnostic.rds\n")
```
