---
title: "Phase 1c: Corpus Refinement for Destination Framing Analysis"
subtitle: "Word boundary disambiguation plus uniform register gate"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(stringi)
library(ggplot2)
library(knitr)
library(patchwork)

for (d in c("data", "data-raw", "output/tables", "output/figures",
            "output/reports")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 1b diagnostics revealed two distinct problems with the raw retrieval
corpus.

1. Polysemic destinations (Sinj, Nin, Solin) suffered substring contamination
   because ILIKE patterns without word boundaries matched inside unrelated
   words (nekretnina, trening, Nina as a given name, sinji as adjective,
   Solinjanin as surname).
2. Even after disambiguation, many destination mentions do not pertain to
   the religious tourism discourse at all (traffic reports, property ads,
   local sports coverage), which inflates the corpus with content outside
   the definitional scope of the study.

Phase 1c applies two sequential refinements to the Phase 1 working corpus
without rerunning database retrieval.

Refinement 1, Unicode aware word boundary filter for polysemic destinations.
Applies only to Sinj, Nin, Solin. Uses stringi regex with proper Unicode
letter class so Croatian diacritics are respected.

Refinement 2, uniform register gate. Article-destination pair is kept only
if at least one of (religious, tourism) registers is present in the full
text or title. Scandal register alone does not qualify as in scope because
scandal content without religious or tourism anchor is outside study scope.

The refinement is applied to all destinations uniformly to preserve one
consistent scope across the analytical corpus. Clean destinations
(Marija Bistrica, Trsat, Aljmaš, Voćin, Vepric, Ludbreg, Krasno) lose
a portion of their volume, but that portion is precisely the part of their
coverage that sits outside the religious tourism frame.

The Phase 1 corpus (destinations_long.rds) is preserved as the provenance
record. The Phase 1c output (destinations_analytical.rds) is the working
corpus for Phases 2 onward.

# 0. Configuration

```{r}
#| label: config

config_c <- list(
  timestamp              = Sys.time(),
  input_corpus           = "data-raw/destinations_long.rds",
  output_corpus          = "data-raw/destinations_analytical.rds",
  output_metadata        = "data/destinations_analytical_metadata.rds",
  polysemic_destinations = c("sinj", "nin", "solin"),
  clean_destinations     = c("marija_bistrica", "trsat", "aljmas",
                             "vocin", "vepric", "ludbreg", "krasno"),
  register_gate_mode     = "uniform",
  register_gate_requires = c("has_religious_register", "has_tourism_register"),
  min_articles_stm       = 300L,
  min_articles_desc      = 50L,
  minimum_months         = 41L,
  max_clean_loss_pct     = 70,
  min_polysemic_kept     = 1000L
)

saveRDS(config_c, "data/config_phase1c.rds")
cat("Phase 1c configuration saved.\n")
cat("  Polysemic destinations:",
    paste(config_c$polysemic_destinations, collapse = ", "), "\n")
cat("  Register gate mode:", config_c$register_gate_mode, "\n")
```

# 1. Load Phase 1 Corpus

```{r}
#| label: load

long <- readRDS(config_c$input_corpus)
setDT(long)

cat("Phase 1 corpus loaded.\n")
cat("  Rows (pairs):", format(nrow(long), big.mark = ","), "\n")
cat("  Unique articles:", format(uniqueN(long$article_id), big.mark = ","), "\n")
cat("  Destinations:", length(unique(long$destination)), "\n\n")

baseline_counts <- long[, .(baseline_pairs = .N), by = destination]
kable(baseline_counts[order(-baseline_pairs)],
      format.args = list(big.mark = ","),
      caption = "Baseline pair counts per destination (Phase 1 input)")
```

# 2. Refinement 1: Word Boundary Disambiguation

Apply Unicode aware regex with letter class boundaries for polysemic
destinations. The pattern structure is as follows.

- Leading guard is start of string or non letter character, expressed as
  `(?:^|[^\\p{L}])`.
- Stem is matched case insensitively.
- Optional suffix is any sequence of letters, expressed as `[\\p{L}]*`.
- Trailing guard is end of string or non letter character, expressed as
  `(?:$|[^\\p{L}])`.

This respects Croatian diacritics because `\p{L}` is the Unicode letter
class, not ASCII `[a-z]`.

```{r}
#| label: boundary-filter

# stem definitions per polysemic destination
# note: Sinj stems deliberately exclude "alka"/"alkar" because those
# introduce event-level content that is separately tracked and would
# otherwise cause pattern-trigger overlap (see Phase 1 review)
boundary_stems <- list(
  sinj  = c("Sinj", "sinjsk"),
  nin   = c("Nin", "ninsk"),
  solin = c("Solin", "solinsk")
)

make_boundary_rx <- function(stems) {
  # escape any regex metachars in stems (defensive, stems should be plain)
  esc <- stri_replace_all_regex(stems, "([.\\\\+*?\\[\\]{}^$()|])", "\\\\$1")
  paste0("(?i)(?:^|[^\\p{L}])(?:",
         paste(esc, collapse = "|"),
         ")[\\p{L}]*(?:$|[^\\p{L}])")
}

# initialise column; non-polysemic rows stay TRUE by default
long[, pass_boundary := TRUE]

# apply boundary regex only to polysemic destinations
# note on residual person-name contamination (Nina, Nino etc.)
# the stem "Nin" with letter-class suffix cannot be distinguished from
# Croatian given names through regex alone. the register gate in step 4
# filters the remainder. empirically ~9 percent of Nin rows are
# person-name matches (per Phase 1b contamination check), most of which
# lack religious or tourism register and drop out at the gate.

for (d in config_c$polysemic_destinations) {
  rx <- make_boundary_rx(boundary_stems[[d]])

  long[destination == d,
       pass_boundary := {
         blob <- paste(TITLE, MENTION_SNIPPET, sep = " | ")
         hit  <- stri_detect_regex(blob, rx)
         hit[is.na(hit)] <- FALSE
         hit
       }]

  n_before <- long[destination == d, .N]
  n_after  <- long[destination == d & pass_boundary == TRUE, .N]
  cat(sprintf("  %-8s | before %7s | after %7s | dropped %5.1f%%\n",
              d,
              format(n_before, big.mark = ","),
              format(n_after, big.mark = ","),
              100 * (1 - n_after / n_before)))
}

cat("\nWord boundary filter applied.\n")
cat("  Clean destinations untouched (pass_boundary = TRUE for all).\n")
```

# 3. Intermediate Corpus After Refinement 1

```{r}
#| label: intermediate-A

long_A <- long[pass_boundary == TRUE]

cat("Corpus after word boundary filter:\n")
cat("  Rows:", format(nrow(long_A), big.mark = ","), "\n")
cat("  Unique articles:", format(uniqueN(long_A$article_id), big.mark = ","), "\n\n")

interim_counts <- long_A[, .(after_A = .N), by = destination]
counts_A <- merge(baseline_counts, interim_counts,
                   by = "destination", all.x = TRUE)
counts_A[is.na(after_A), after_A := 0]
counts_A[, pct_kept_A := round(after_A / baseline_pairs * 100, 1)]
counts_A <- counts_A[order(-baseline_pairs)]

kable(counts_A, format.args = list(big.mark = ","),
      caption = "After word boundary filter")
```

# 4. Refinement 2: Uniform Register Gate

Keep pair if religious or tourism register is present. Applies to all
destinations, not only polysemic ones. This enforces a single scope
across the analytical corpus.

```{r}
#| label: register-gate

long_A[, pass_register := has_religious_register | has_tourism_register]
long_A[is.na(pass_register), pass_register := FALSE]

# diagnostic before we filter
register_gate_summary <- long_A[, .(
  n = .N,
  pass_register = sum(pass_register),
  pct_kept = round(sum(pass_register) / .N * 100, 1)
), by = destination][order(-n)]

kable(register_gate_summary, format.args = list(big.mark = ","),
      caption = "Register gate retention rate by destination")
```

# 5. Final Analytical Corpus

```{r}
#| label: final-corpus

long_final <- long_A[pass_register == TRUE]

cat("Final analytical corpus:\n")
cat("  Rows:", format(nrow(long_final), big.mark = ","), "\n")
cat("  Unique articles:", format(uniqueN(long_final$article_id),
                                 big.mark = ","), "\n")
cat("  Months covered:", uniqueN(long_final$yearmonth),
    "of", config_c$minimum_months, "\n\n")

final_counts <- long_final[, .(final_pairs = .N), by = destination]
all_counts <- merge(counts_A, final_counts, by = "destination",
                     all.x = TRUE)
all_counts[is.na(final_pairs), final_pairs := 0]
all_counts[, pct_kept_total := round(final_pairs / baseline_pairs * 100, 1)]
all_counts[, pct_kept_given_A := round(final_pairs / after_A * 100, 1)]
all_counts <- all_counts[order(-baseline_pairs)]

kable(all_counts, format.args = list(big.mark = ","),
      caption = "Pair counts across all refinement stages")
```

# 6. Sanity Checks on Clean Destinations

The clean destinations (Marija Bistrica, Trsat, Aljmaš, Voćin, Vepric,
Ludbreg, Krasno) should retain most of their volume through the register
gate. If any of them loses more than the configured threshold, that
indicates either a register dictionary gap or an unexpected content
profile that warrants investigation.

```{r}
#| label: sanity-clean

clean_check <- all_counts[destination %in% config_c$clean_destinations,
                           .(destination, baseline_pairs, final_pairs,
                             pct_kept_total)]
clean_check[, status := fifelse(
  pct_kept_total >= (100 - config_c$max_clean_loss_pct),
  "OK", "REVIEW")]

kable(clean_check, format.args = list(big.mark = ","),
      caption = "Clean destinations sanity check")

any_review <- any(clean_check$status == "REVIEW")
if (any_review) {
  cat("\nNOTE: at least one clean destination lost more than",
      config_c$max_clean_loss_pct, "percent.\n")
  cat("Inspect sample of dropped articles before proceeding.\n")
} else {
  cat("\nAll clean destinations retained at least",
      100 - config_c$max_clean_loss_pct, "percent. Proceed.\n")
}
```

# 7. Sanity Checks on Polysemic Destinations

After both refinements, polysemic destinations should sit closer to the
clean destination baseline in terms of register prevalence. They should
also retain enough absolute volume for downstream STM estimation.

```{r}
#| label: sanity-polysemic

poly_check <- long_final[destination %in% config_c$polysemic_destinations,
  .(
    final_pairs = .N,
    pct_religious = round(mean(has_religious_register) * 100, 1),
    pct_tourism = round(mean(has_tourism_register) * 100, 1),
    pct_any_target = round(mean(has_religious_register |
                                 has_tourism_register) * 100, 1)
  ), by = destination]

clean_reference <- long_final[destination %in% config_c$clean_destinations,
  .(
    final_pairs = .N,
    pct_religious = round(mean(has_religious_register) * 100, 1),
    pct_tourism = round(mean(has_tourism_register) * 100, 1),
    pct_any_target = round(mean(has_religious_register |
                                 has_tourism_register) * 100, 1)
  ), by = destination]

poly_check[, status := fifelse(
  final_pairs >= config_c$min_polysemic_kept, "OK", "REVIEW")]

kable(rbind(
  cbind(group = "polysemic", poly_check),
  cbind(group = "clean", clean_reference[, .(
    destination, final_pairs, pct_religious, pct_tourism, pct_any_target,
    status = "reference")])
), format.args = list(big.mark = ","),
   caption = "Polysemic vs clean destinations after full refinement")
```

# 8. Frame Variable Construction

With the analytical corpus defined, construct the dominant frame variable
that operationalises the second research question (context of media
mentions). The dominant frame is assigned according to a hierarchy that
respects the research definition.

Priority order:
1. Scandal frame if scandal register present without religious anchor.
2. Religious tourism mixed frame if both religious and tourism registers
   present.
3. Religious frame if religious register only.
4. Tourism frame if tourism register only.
5. Other, residual category.

```{r}
#| label: frame-construction

long_final[, dominant_frame := fcase(
  has_scandal_register & !has_religious_register & !has_tourism_register,
    "scandal",
  has_religious_register & has_tourism_register,
    "religious_tourism_mixed",
  has_religious_register & !has_tourism_register & has_scandal_register,
    "religious_with_scandal",
  has_religious_register & !has_tourism_register,
    "religious_only",
  has_tourism_register & !has_religious_register & has_scandal_register,
    "tourism_with_scandal",
  has_tourism_register & !has_religious_register,
    "tourism_only",
  default = "other"
)]

frame_dist <- long_final[, .N, by = .(destination, dominant_frame)]
frame_dist_wide <- dcast(frame_dist, destination ~ dominant_frame,
                          value.var = "N", fill = 0L)

# total per destination
frame_dist_wide[, total := rowSums(.SD),
                .SDcols = setdiff(names(frame_dist_wide), "destination")]
frame_dist_wide <- frame_dist_wide[order(-total)]

kable(frame_dist_wide, format.args = list(big.mark = ","),
      caption = "Dominant frame distribution per destination (counts)")

# same as percentages
frame_cols <- setdiff(names(frame_dist_wide), c("destination", "total"))
frame_pct <- copy(frame_dist_wide)
for (col in frame_cols) {
  frame_pct[, (col) := round(get(col) / total * 100, 1)]
}
frame_pct[, total := NULL]

kable(frame_pct,
      caption = "Dominant frame distribution per destination (percentages)")
```

# 9. Visual Diagnostics

```{r}
#| label: visual-retention

retention_long <- melt(all_counts[, .(destination, baseline_pairs,
                                       after_A, final_pairs)],
                       id.vars = "destination",
                       variable.name = "stage",
                       value.name = "pairs")
retention_long[, stage := factor(stage,
                                  levels = c("baseline_pairs",
                                             "after_A",
                                             "final_pairs"),
                                  labels = c("Phase 1",
                                             "After boundary filter",
                                             "After register gate"))]

p_retention <- ggplot(retention_long,
                      aes(stage, pairs, group = destination,
                          colour = destination)) +
  geom_line(alpha = 0.7) +
  geom_point(size = 2) +
  scale_y_log10(labels = scales::comma) +
  labs(title = "Pair count retention across refinement stages",
       subtitle = "Log scale. Steep drops on sinj and nin are intentional.",
       x = NULL, y = "Pairs (log scale)") +
  theme_minimal() +
  theme(legend.position = "right")

ggsave("output/figures/01c_retention_by_stage.png", p_retention,
       width = 10, height = 6)
p_retention
```

```{r}
#| label: visual-frames

frame_plot_data <- long_final[, .N, by = .(destination, dominant_frame)]
frame_plot_data[, total := sum(N), by = destination]
frame_plot_data[, pct := N / total * 100]
frame_plot_data[, destination := factor(destination,
                                         levels = all_counts[order(-final_pairs),
                                                              destination])]

p_frames <- ggplot(frame_plot_data,
                   aes(destination, pct, fill = dominant_frame)) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Dominant frame composition by destination",
       x = NULL, y = "Percent of pairs", fill = "Frame") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "bottom")

ggsave("output/figures/01c_frame_composition.png", p_frames,
       width = 12, height = 6)
p_frames
```

```{r}
#| label: visual-monthly

monthly_final <- long_final[, .N, by = .(destination, yearmonth)]
p_monthly <- ggplot(monthly_final, aes(yearmonth, N, colour = destination)) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ destination, scales = "free_y", ncol = 2) +
  labs(title = "Monthly volume by destination (analytical corpus)",
       x = NULL, y = "Pairs") +
  theme_minimal() +
  theme(legend.position = "none")

ggsave("output/figures/01c_monthly_volume_analytical.png", p_monthly,
       width = 12, height = 10)
p_monthly
```

# 10. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 1c\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. all destinations retained at least descriptive minimum
desc_check <- long_final[, .N, by = destination]
desc_check[, pass := N >= config_c$min_articles_desc]
checks$all_above_descriptive <- all(desc_check$pass)
cat("1. All destinations above ", config_c$min_articles_desc, " pairs: ",
    ifelse(checks$all_above_descriptive, "PASS", "CAUTION"), "\n", sep = "")
if (!checks$all_above_descriptive) {
  thin <- desc_check[pass == FALSE]$destination
  cat("   Below threshold:", paste(thin, collapse = ", "), "\n")
}

# 2. at least N destinations above STM threshold
n_above_stm <- sum(desc_check$N >= config_c$min_articles_stm)
checks$stm_ready <- n_above_stm >= 6
cat("2. At least 6 destinations above ", config_c$min_articles_stm,
    " (STM-ready): ",
    ifelse(checks$stm_ready, "PASS", "CAUTION"),
    " (", n_above_stm, ")\n", sep = "")

# 3. clean destinations retained sufficient volume
clean_retention <- all_counts[destination %in% config_c$clean_destinations]
checks$clean_retained <- all(clean_retention$pct_kept_total >=
                              (100 - config_c$max_clean_loss_pct))
cat("3. Clean destinations retained at least ",
    100 - config_c$max_clean_loss_pct, "%: ",
    ifelse(checks$clean_retained, "PASS", "REVIEW"), "\n", sep = "")

# 4. polysemic destinations retained minimum absolute volume
poly_volume <- long_final[destination %in% config_c$polysemic_destinations,
                           .N, by = destination]
checks$poly_volume <- all(poly_volume$N >= config_c$min_polysemic_kept)
cat("4. Polysemic destinations above ",
    format(config_c$min_polysemic_kept, big.mark = ","), " pairs: ",
    ifelse(checks$poly_volume, "PASS", "CAUTION"), "\n", sep = "")
if (!checks$poly_volume) {
  thin <- poly_volume[N < config_c$min_polysemic_kept]
  cat("   Below threshold: ",
      paste0(thin$destination, " (", format(thin$N, big.mark = ","), ")",
             collapse = ", "),
      "\n")
}

# 5. register prevalence gap between polysemic and clean narrowed
poly_any <- long_final[destination %in% config_c$polysemic_destinations,
                        mean(has_religious_register |
                             has_tourism_register)]
clean_any <- long_final[destination %in% config_c$clean_destinations,
                         mean(has_religious_register |
                              has_tourism_register)]
# both should be 100% by construction after register gate
# sanity check, both should equal 1
checks$gate_enforced <- isTRUE(all.equal(poly_any, 1)) &&
                         isTRUE(all.equal(clean_any, 1))
cat("5. Register gate correctly enforced (both groups at 100%): ",
    ifelse(checks$gate_enforced, "PASS", "FAIL"),
    " (poly=", round(poly_any * 100, 1),
    "%, clean=", round(clean_any * 100, 1), "%)\n", sep = "")

# 6. months covered
checks$months_full <- uniqueN(long_final$yearmonth) >= config_c$minimum_months
cat("6. All ", config_c$minimum_months, " months covered: ",
    ifelse(checks$months_full, "PASS", "FAIL"),
    " (", uniqueN(long_final$yearmonth), ")\n", sep = "")

# 7. frame diversity per major destination
major_dests <- desc_check[N >= config_c$min_articles_stm]$destination
frame_div <- long_final[destination %in% major_dests,
                         .(n_frames = uniqueN(dominant_frame)),
                         by = destination]
checks$frame_diversity <- all(frame_div$n_frames >= 3)
cat("7. Major destinations span at least 3 frame categories: ",
    ifelse(checks$frame_diversity, "PASS", "NOTE"), "\n", sep = "")

cat("\n")
n_pass <- sum(unlist(checks))
cat("Passed:", n_pass, "of", length(checks), "\n")
if (all(unlist(checks))) {
  cat("ALL CHECKS PASSED. destinations_analytical.rds ready for Phase 2.\n")
} else {
  failed <- names(checks)[!unlist(checks)]
  cat("Attention needed:", paste(failed, collapse = ", "), "\n")
}
cat(strrep("=", 60), "\n")
```

# 11. Save Outputs

```{r}
#| label: save

# drop helper columns before saving the analytical corpus
long_final[, pass_boundary := NULL]
long_final[, pass_register := NULL]

saveRDS(long_final, config_c$output_corpus)

refinement_meta <- list(
  timestamp        = Sys.time(),
  phase            = "1c_corpus_refinement",
  input_corpus     = config_c$input_corpus,
  output_corpus    = config_c$output_corpus,
  r_version        = R.version.string,
  config           = config_c,
  boundary_stems   = boundary_stems,
  baseline_counts  = baseline_counts,
  counts_after_A   = counts_A,
  final_counts     = final_counts,
  all_counts       = all_counts,
  frame_dist_wide  = frame_dist_wide,
  frame_pct        = frame_pct,
  checks           = checks,
  n_pairs_input    = nrow(long),
  n_pairs_after_A  = nrow(long_A),
  n_pairs_final    = nrow(long_final),
  n_articles_final = uniqueN(long_final$article_id),
  months_covered   = uniqueN(long_final$yearmonth)
)

saveRDS(refinement_meta, config_c$output_metadata)

fwrite(all_counts, "output/tables/01c_retention_table.csv")
fwrite(frame_dist_wide, "output/tables/01c_frame_distribution_counts.csv")
fwrite(frame_pct, "output/tables/01c_frame_distribution_pct.csv")

cat("Outputs saved.\n")
cat("  Analytical corpus:", config_c$output_corpus, "\n")
cat("  Metadata:        ", config_c$output_metadata, "\n\n")

fs <- c(config_c$output_corpus, config_c$output_metadata,
        "output/tables/01c_retention_table.csv",
        "output/tables/01c_frame_distribution_counts.csv",
        "output/tables/01c_frame_distribution_pct.csv",
        "output/figures/01c_retention_by_stage.png",
        "output/figures/01c_frame_composition.png",
        "output/figures/01c_monthly_volume_analytical.png")
for (f in fs) {
  if (file.exists(f)) {
    size_mb <- round(file.size(f) / 1e6, 2)
    cat("  ", f, " | ", size_mb, " MB\n", sep = "")
  }
}
```

# 12. Next Steps

The analytical corpus destinations_analytical.rds now carries one
consistent definitional scope. Phase 2 can proceed along three tracks.

First, descriptive profile. Frequency of each destination, platform
distribution, source concentration (is coverage driven by few outlets),
and syndication rate estimated via content hash clustering.

Second, temporal dynamics. Monthly volume decomposed into trend, seasonal
cycle, and residual. Overlay of liturgical calendar events (Velika Gospa,
Mala Gospa, patronal feasts per destination) on the monthly series to test
whether seasonality follows feast days or is uniformly distributed.

Third, framing analysis. Dominant frame as categorical outcome, with
destination and month as predictors. STM on the full text corpus with
prevalence covariates destination, month, and source type.

Phase 1c has closed the retrieval stage. Phase 2 is descriptive analysis
on the settled analytical corpus.
