---
title: "Phase 2b: Temporal Dynamics and Seasonality"
subtitle: "Monthly decomposition, liturgical calendar overlay, seasonality test"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(ggplot2)
library(knitr)
library(patchwork)
library(forecast)

for (d in c("data", "output/tables", "output/figures", "output/reports")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 2b addresses the third research question: whether media coverage of
religious tourism destinations clusters around liturgical events, or is
distributed uniformly across the calendar.

Four analytical components.

1. Monthly volume time series per destination with trend and seasonal
   components extracted via STL decomposition.
2. Liturgical calendar map per destination, anchoring each destination to
   its principal patronal feasts.
3. Feast proximity analysis. For each destination, articles are labelled
   by their distance to the nearest feast day. The distribution of
   mentions across distance bins reveals whether coverage is feast
   anchored or continuous.
4. Seasonality index per destination, defined as the ratio of feast month
   volume to non feast month volume. A large ratio signals event driven
   coverage, a ratio near one signals continuous coverage.

The analytical consequence for the research question is straightforward.
Event anchored destinations function in the media as calendar ritual
sites. Continuously covered destinations function as standing tourist
destinations. This distinction is the empirical core of the third
research question.

# 0. Configuration

```{r}
#| label: config

config_2b <- list(
  timestamp        = Sys.time(),
  input_corpus     = "data/destinations_analytical_cleaned.rds",
  output_metadata  = "data/phase2b_metadata.rds",
  feast_window_days = 14L,
  dedup_by_cluster = FALSE,
  min_months_for_stl = 24L
)
saveRDS(config_2b, "data/config_phase2b.rds")
cat("Phase 2b configuration saved.\n")
```

# 1. Load Enriched Corpus

```{r}
#| label: load

# fall back to plain analytical corpus if enriched is not available
input_path <- config_2b$input_corpus
if (!file.exists(input_path)) {
  input_path <- "../01-retrieval/data-raw/destinations_analytical_clean.rds"
  cat("Note: using plain analytical corpus, syndication weighting disabled.\n")
}

long <- readRDS(input_path)
setDT(long)
long <- long[cleaning_action == "keep"]

if (!"yearmonth" %in% names(long))
  long[, yearmonth := as.Date(format(date, "%Y-%m-01"))]
if (!"cluster_size" %in% names(long)) long[, cluster_size := 1L]

cat("Corpus loaded.\n")
cat("  Rows:", format(nrow(long), big.mark = ","), "\n")
cat("  Date range:", format(min(long$date)), "to",
    format(max(long$date)), "\n\n")
```

# 2. Liturgical Calendar

Feast dates that define the patronal calendar for each destination.
Fixed date feasts are stored as month-day pairs. Movable feasts are
excluded from the core overlay because they do not align reliably
across years and their effect is better studied separately.

```{r}
#| label: liturgical-calendar

# each entry is a list of feasts; a feast has a name and month-day
liturgical_calendar <- list(

  marija_bistrica = list(
    list(name = "Velika Gospa",    mm = 8L, dd = 15L),
    list(name = "Mala Gospa",      mm = 9L, dd = 8L),
    list(name = "Rođenje Marijino u Bistrici", mm = 7L, dd = 13L)
  ),

  trsat = list(
    list(name = "Translatio Trsat",  mm = 5L, dd = 10L),
    list(name = "Velika Gospa",      mm = 8L, dd = 15L),
    list(name = "Mala Gospa",        mm = 9L, dd = 8L)
  ),

  sinj = list(
    list(name = "Velika Gospa Sinjska", mm = 8L, dd = 15L),
    list(name = "Sinjska Alka",         mm = 8L, dd = 7L),
    list(name = "Hodočasnička devetnica", mm = 8L, dd = 6L)
  ),

  aljmas = list(
    list(name = "Mala Gospa",        mm = 9L, dd = 8L),
    list(name = "Velika Gospa",      mm = 8L, dd = 15L)
  ),

  ludbreg = list(
    list(name = "Sveta Nedjelja ludbreška", mm = 9L, dd = 1L),
    list(name = "Svetkovina Presvete Krvi", mm = 9L, dd = 2L),
    list(name = "Gospin rođendan (Mala Gospa)", mm = 9L, dd = 8L)
  ),

  krasno = list(
    list(name = "Gospa od Krasna",   mm = 8L, dd = 15L),
    list(name = "Svetkovina Krasnoga",mm = 7L, dd = 2L)
  ),

  vocin = list(
    list(name = "Velika Gospa Voćinska", mm = 8L, dd = 15L),
    list(name = "Mala Gospa",            mm = 9L, dd = 8L)
  ),

  vepric = list(
    list(name = "Gospa Lurdska",     mm = 2L, dd = 11L),
    list(name = "Velika Gospa",      mm = 8L, dd = 15L)
  ),

  nin = list(
    list(name = "Gospa od Zečeva (svibanj)", mm = 5L, dd = 5L),
    list(name = "Gospa od Zečeva (kolovoz)", mm = 8L, dd = 5L),
    list(name = "Sveti Nikola Tavelić",      mm = 11L, dd = 14L)
  ),

  solin = list(
    list(name = "Gospa od Otoka",    mm = 9L, dd = 12L),
    list(name = "Gospa od Otoka (glavna)", mm = 9L, dd = 13L),
    list(name = "Mala Gospa",        mm = 9L, dd = 8L)
  )
)

# unpack to long data.table
feast_dt <- rbindlist(lapply(names(liturgical_calendar), function(d) {
  feasts <- liturgical_calendar[[d]]
  rbindlist(lapply(feasts, function(f) {
    data.table(destination = d, feast_name = f$name,
                mm = f$mm, dd = f$dd)
  }))
}))

# expand to actual calendar dates across analysis years
years <- sort(unique(year(long$date)))
feast_dates <- rbindlist(lapply(years, function(y) {
  feast_dt[, .(destination, feast_name, mm, dd,
                feast_date = as.Date(sprintf("%d-%02d-%02d", y, mm, dd)))]
}))
feast_dates <- feast_dates[feast_date >= min(long$date) &
                             feast_date <= max(long$date)]

kable(feast_dt, caption = "Patronal feast calendar per destination")
fwrite(feast_dt, "output/tables/02b_feast_calendar.csv")
fwrite(feast_dates, "output/tables/02b_feast_dates_expanded.csv")
```

# 3. Monthly Volume Time Series

```{r}
#| label: monthly-volume

monthly_ts <- long[, .(pairs = .N), by = .(destination, yearmonth)]

# ensure dense calendar (fill zero months)
all_months <- seq(min(long$yearmonth), max(long$yearmonth), by = "month")
scaffold <- CJ(destination = unique(long$destination),
                yearmonth = all_months)
monthly_ts <- merge(scaffold, monthly_ts,
                     by = c("destination", "yearmonth"), all.x = TRUE)
monthly_ts[is.na(pairs), pairs := 0L]

p_monthly <- ggplot(monthly_ts, aes(yearmonth, pairs)) +
  geom_col(fill = "#2E4057") +
  facet_wrap(~ destination, scales = "free_y", ncol = 2) +
  labs(title = "Monthly volume by destination",
       x = NULL, y = "Pairs per month") +
  theme_minimal()

ggsave("output/figures/02b_monthly_volume.png", p_monthly,
       width = 12, height = 10)
p_monthly

fwrite(monthly_ts, "output/tables/02b_monthly_volume.csv")
```

# 4. STL Decomposition per Destination

```{r}
#| label: stl-decomposition

# only run STL where enough data
monthly_counts_check <- monthly_ts[, .(n_months = .N,
                                         non_zero = sum(pairs > 0)),
                                    by = destination]
stl_eligible <- monthly_counts_check[
  non_zero >= config_2b$min_months_for_stl]$destination

cat("STL eligible destinations (", length(stl_eligible), "):\n")
cat("  ", paste(stl_eligible, collapse = ", "), "\n\n")

stl_results <- list()

for (d in stl_eligible) {
  series <- monthly_ts[destination == d][order(yearmonth)]

  # frequency 12 for monthly seasonality
  ts_vec <- ts(series$pairs,
                start = c(year(min(series$yearmonth)),
                          month(min(series$yearmonth))),
                frequency = 12)

  decomp <- tryCatch(
    stl(ts_vec, s.window = "periodic", robust = TRUE),
    error = function(e) NULL
  )

  if (is.null(decomp)) {
    cat("  ", d, " STL failed, skipping\n", sep = "")
    next
  }

  comp <- as.data.table(decomp$time.series)
  comp[, yearmonth := series$yearmonth]
  comp[, destination := d]
  comp[, pairs := series$pairs]

  stl_results[[d]] <- comp
}

stl_all <- rbindlist(stl_results)

# strength of seasonality and strength of trend, following Hyndman
season_strength <- stl_all[, {
  var_resid <- var(remainder)
  var_resid_plus_season <- var(remainder + seasonal)
  var_resid_plus_trend  <- var(remainder + trend)
  .(strength_seasonal = round(
      max(0, 1 - var_resid / var_resid_plus_season), 3),
    strength_trend = round(
      max(0, 1 - var_resid / var_resid_plus_trend), 3))
}, by = destination][order(-strength_seasonal)]

kable(season_strength,
      caption = "Strength of seasonal and trend components (Hyndman & Athanasopoulos)")

fwrite(season_strength, "output/tables/02b_season_strength.csv")
```

```{r}
#| label: stl-plots

for (d in stl_eligible) {
  comp <- stl_results[[d]]
  if (is.null(comp)) next

  p_data <- rbind(
    comp[, .(yearmonth, value = pairs,     component = "observed")],
    comp[, .(yearmonth, value = trend,     component = "trend")],
    comp[, .(yearmonth, value = seasonal,  component = "seasonal")],
    comp[, .(yearmonth, value = remainder, component = "remainder")]
  )
  p_data[, component := factor(component,
                                levels = c("observed", "trend",
                                           "seasonal", "remainder"))]

  p <- ggplot(p_data, aes(yearmonth, value)) +
    geom_line(colour = "#2E4057") +
    facet_wrap(~ component, ncol = 1, scales = "free_y") +
    labs(title = paste0("STL decomposition: ", d),
         x = NULL, y = NULL) +
    theme_minimal()

  ggsave(paste0("output/figures/02b_stl_", d, ".png"),
         p, width = 10, height = 8)
}
cat("STL plots saved to output/figures/\n")
```

# 5. Feast Proximity Analysis

For each article, compute the minimum distance in days to the nearest
patronal feast of its own destination. Bin by week of proximity.

```{r}
#| label: feast-proximity

feast_by_dest <- split(feast_dates, feast_dates$destination)

compute_distance <- function(dest, date) {
  fd <- feast_by_dest[[dest]]$feast_date
  if (is.null(fd) || length(fd) == 0) return(NA_integer_)
  min(abs(as.integer(date - fd)))
}

# vectorised for speed
long[, dist_to_feast := NA_integer_]

for (d in names(feast_by_dest)) {
  fd <- feast_by_dest[[d]]$feast_date
  idx <- which(long$destination == d)
  if (length(idx) == 0) next

  dists <- sapply(long$date[idx], function(x) min(abs(as.integer(x - fd))))
  long[idx, dist_to_feast := dists]
}

# bin: 0-7 (feast week), 8-14 (next week), 15-30 (feast month), 31-60,
# 61-120, 121-180, 181+ (off season)
long[, feast_bin := cut(dist_to_feast,
                         breaks = c(-Inf, 7, 14, 30, 60, 120, 180, Inf),
                         labels = c("0-7", "8-14", "15-30", "31-60",
                                    "61-120", "121-180", "181+"),
                         right = TRUE)]

bin_dist <- long[!is.na(feast_bin),
                  .N, by = .(destination, feast_bin)]
bin_dist[, total := sum(N), by = destination]
bin_dist[, pct := round(N / total * 100, 1)]

bin_wide <- dcast(bin_dist, destination ~ feast_bin,
                   value.var = "pct", fill = 0)
kable(bin_wide, caption = "Share of pairs by proximity to nearest patronal feast (%)")

fwrite(bin_wide, "output/tables/02b_feast_proximity_pct.csv")
```

```{r}
#| label: feast-proximity-plot

bin_dist[, destination := factor(destination,
                                   levels = bin_wide[order(-`0-7`),
                                                      destination])]

p_prox <- ggplot(bin_dist, aes(feast_bin, pct, fill = destination)) +
  geom_col(position = "dodge") +
  facet_wrap(~ destination, scales = "free_y", ncol = 2) +
  labs(title = "Distance from patronal feast",
       subtitle = "High share in 0-7 bucket indicates feast anchored coverage",
       x = "Days to nearest feast", y = "Percent of pairs") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1))

ggsave("output/figures/02b_feast_proximity.png", p_prox,
       width = 12, height = 10)
p_prox
```

# 6. Feast Window Volume vs Baseline

```{r}
#| label: feast-window

long[, in_feast_window := dist_to_feast <= config_2b$feast_window_days]
long[is.na(in_feast_window), in_feast_window := FALSE]

window_analysis <- long[, .(
  total_pairs = .N,
  window_pairs = sum(in_feast_window),
  window_share = round(sum(in_feast_window) / .N * 100, 1)
), by = destination]

# compute how many days fall inside any window per destination
# total days in analysis window
total_days <- as.integer(max(long$date) - min(long$date)) + 1

# days inside feast windows per destination
window_days_per_dest <- sapply(names(feast_by_dest), function(d) {
  fd <- feast_by_dest[[d]]$feast_date
  if (length(fd) == 0) return(0)
  # all days within feast_window_days of any feast
  all_d <- seq(min(long$date), max(long$date), by = "day")
  dists <- sapply(all_d, function(x) min(abs(as.integer(x - fd))))
  sum(dists <= config_2b$feast_window_days)
})

window_analysis[, days_in_window := window_days_per_dest[destination]]
window_analysis[, expected_share_if_uniform :=
                  round(days_in_window / total_days * 100, 1)]
window_analysis[, seasonality_index :=
                  round(window_share / expected_share_if_uniform, 2)]
window_analysis <- window_analysis[order(-seasonality_index)]

kable(window_analysis, format.args = list(big.mark = ","),
      caption = paste0("Feast window analysis, ",
                       config_2b$feast_window_days, " day window each side"))

fwrite(window_analysis, "output/tables/02b_feast_window_analysis.csv")
```

```{r}
#| label: seasonality-interpretation

# interpret seasonality index
window_analysis[, pattern := fcase(
  seasonality_index >= 3.0, "heavily feast anchored",
  seasonality_index >= 1.8, "moderately feast anchored",
  seasonality_index >= 1.2, "mildly feast anchored",
  seasonality_index >= 0.8, "continuous or uniform",
  default = "counter-cyclical or off-season heavy"
)]

kable(window_analysis[, .(destination, window_share,
                           expected_share_if_uniform,
                           seasonality_index, pattern)],
      caption = "Seasonality interpretation per destination")
```

# 7. Chi Squared Test of Feast vs Non Feast Distribution

Test whether observed feast window volume differs significantly from
expected under uniform temporal distribution.

```{r}
#| label: chi-square-test

chi_results <- data.table(destination = character(),
                           observed_window = integer(),
                           observed_off = integer(),
                           expected_window = numeric(),
                           expected_off = numeric(),
                           chi_sq = numeric(),
                           p_value = numeric())

for (d in unique(long$destination)) {
  dest_data <- long[destination == d]
  if (nrow(dest_data) < 30) next

  observed_w <- sum(dest_data$in_feast_window)
  observed_o <- sum(!dest_data$in_feast_window)

  prop_w <- window_days_per_dest[d] / total_days
  expected_w <- nrow(dest_data) * prop_w
  expected_o <- nrow(dest_data) * (1 - prop_w)

  if (expected_w < 5 || expected_o < 5) next

  chi <- chisq.test(x = c(observed_w, observed_o),
                     p = c(prop_w, 1 - prop_w))

  chi_results <- rbind(chi_results, data.table(
    destination = d,
    observed_window = observed_w,
    observed_off = observed_o,
    expected_window = round(expected_w, 1),
    expected_off = round(expected_o, 1),
    chi_sq = round(unname(chi$statistic), 2),
    p_value = chi$p.value
  ))
}

chi_results[, significant := p_value < 0.001]
chi_results <- chi_results[order(-chi_sq)]

kable(chi_results, format.args = list(big.mark = ","),
      caption = "Chi squared test: feast window vs uniform distribution")

fwrite(chi_results, "output/tables/02b_chi_square_seasonality.csv")
```

# 8. Frame Composition Inside vs Outside Feast Windows

Beyond pure volume, examine whether the composition of frames changes
between feast windows and off-season. A destination whose religious
frame dominates inside the window but whose tourism frame dominates
outside the window tells a richer story than volume alone.

```{r}
#| label: frame-by-window

if (!"dominant_frame" %in% names(long)) {
  long[, dominant_frame := fcase(
    has_scandal_register & !has_religious_register & !has_tourism_register,
      "scandal",
    has_religious_register & has_tourism_register,
      "religious_tourism_mixed",
    has_religious_register & !has_tourism_register & has_scandal_register,
      "religious_with_scandal",
    has_religious_register & !has_tourism_register,
      "religious_only",
    has_tourism_register & !has_religious_register & has_scandal_register,
      "tourism_with_scandal",
    has_tourism_register & !has_religious_register,
      "tourism_only",
    default = "other"
  )]
}

frame_by_window <- long[, .N,
                         by = .(destination, in_feast_window,
                                 dominant_frame)]
frame_by_window[, total := sum(N), by = .(destination, in_feast_window)]
frame_by_window[, pct := round(N / total * 100, 1)]

kable(dcast(frame_by_window,
             destination + in_feast_window ~ dominant_frame,
             value.var = "pct", fill = 0),
      caption = "Frame composition by feast window (%)")

fwrite(frame_by_window, "output/tables/02b_frame_by_window.csv")
```

```{r}
#| label: frame-shift-plot

frame_by_window[, in_feast_window := factor(
  in_feast_window, levels = c(FALSE, TRUE),
  labels = c("off season", "feast window"))]

p_frame_shift <- ggplot(frame_by_window,
                        aes(in_feast_window, pct, fill = dominant_frame)) +
  geom_col() +
  facet_wrap(~ destination, ncol = 3) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Frame composition: feast window vs off season",
       x = NULL, y = "Percent of pairs", fill = "Frame") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 15, hjust = 1))

ggsave("output/figures/02b_frame_shift_by_window.png", p_frame_shift,
       width = 12, height = 10)
p_frame_shift
```

# 9. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2b\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. STL ran for at least 6 destinations
checks$stl_coverage <- length(stl_results) >= 6
cat("1. STL decomposition completed for 6+ destinations: ",
    ifelse(checks$stl_coverage, "PASS", "CAUTION"),
    " (", length(stl_results), ")\n", sep = "")

# 2. at least half of destinations show significant chi squared result
n_sig <- sum(chi_results$significant, na.rm = TRUE)
checks$chi_coverage <- n_sig >= nrow(chi_results) / 2
cat("2. At least half of destinations show significant chi squared: ",
    ifelse(checks$chi_coverage, "PASS", "NOTE"),
    " (", n_sig, " of ", nrow(chi_results), ")\n", sep = "")

# 3. at least one destination each in feast anchored and continuous groups
has_anchored <- any(window_analysis$seasonality_index >= 1.8)
has_continuous <- any(window_analysis$seasonality_index >= 0.8 &
                       window_analysis$seasonality_index < 1.2)
checks$pattern_diversity <- has_anchored & has_continuous
cat("3. Both feast-anchored and continuous patterns present: ",
    ifelse(checks$pattern_diversity, "PASS", "NOTE"), "\n", sep = "")

# 4. frame shift between window and off season is observable
# look for destinations where religious frame share changes by 10+ points
shift_test <- dcast(frame_by_window[dominant_frame == "religious_only"],
                     destination ~ in_feast_window,
                     value.var = "pct", fill = 0)
if ("off season" %in% names(shift_test) &&
    "feast window" %in% names(shift_test)) {
  shift_test[, shift := abs(`feast window` - `off season`)]
  checks$frame_shift_visible <- any(shift_test$shift >= 10)
  cat("4. At least one destination shows 10+ point frame shift: ",
      ifelse(checks$frame_shift_visible, "PASS", "NOTE"), "\n", sep = "")
} else {
  checks$frame_shift_visible <- FALSE
  cat("4. Frame shift analysis failed, check source data\n")
}

cat("\n")
n_pass <- sum(unlist(checks))
cat("Passed:", n_pass, "of", length(checks), "\n")
if (all(unlist(checks))) {
  cat("ALL CHECKS PASSED. Phase 2c can proceed.\n")
} else {
  failed <- names(checks)[!unlist(checks)]
  cat("Attention needed:", paste(failed, collapse = ", "), "\n")
}
cat(strrep("=", 60), "\n")
```

# 10. Save Outputs

```{r}
#| label: save

meta_2b <- list(
  timestamp          = Sys.time(),
  phase              = "2b_seasonality",
  config             = config_2b,
  liturgical_calendar = liturgical_calendar,
  feast_dt           = feast_dt,
  feast_dates        = feast_dates,
  monthly_ts         = monthly_ts,
  season_strength    = season_strength,
  bin_dist           = bin_dist,
  window_analysis    = window_analysis,
  chi_results        = chi_results,
  frame_by_window    = frame_by_window,
  checks             = checks
)
saveRDS(meta_2b, config_2b$output_metadata)

# save long corpus enriched with feast proximity columns
saveRDS(long, "data/destinations_analytical_with_feast.rds")
cat("Phase 2b metadata and enriched corpus saved.\n")
```

# 11. Interpretive Summary

Phase 2b has produced four interpretive layers that directly answer the
third research question.

First, the STL seasonal strength index quantifies how much of each
destination's volume variation is attributable to a recurring annual
cycle. Values above 0.6 indicate strongly cyclic coverage.

Second, the feast proximity distribution shows how concentrated each
destination's coverage is around patronal dates. The share of pairs in
the 0-7 day bin is the simplest single number indicator of feast
anchoring.

Third, the seasonality index compares observed feast window share to
the share expected under uniform distribution. Values above 1.8 signal
meaningful over representation during feast windows, values near 1
signal continuous coverage.

Fourth, the frame shift analysis reveals whether a destination changes
its media profile during feast windows. Ludbreg may for instance be
almost exclusively religious framed year round, while Trsat may shift
from tourism framing in summer to religious framing around patronal
feasts. This layer is particularly informative for the research
question about how media construct the destination.

Phase 2c will bring these three layers together through a framing model
that includes month and destination as covariates.
