---
title: "Phase 2c: Framing Analysis (Revised)"
subtitle: "Refined scandal register, stable multinomial model, two-dimensional typology"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(stringi)
library(ggplot2)
library(knitr)
library(patchwork)
library(nnet)
library(MASS)

for (d in c("data", "output/tables", "output/figures",
            "output/reports", "data-raw")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 2c produces the framing analysis that directly answers the second
research question (in which context destinations are mentioned) and
integrates Phase 2a and 2b findings into a unified two-dimensional
typology.

This revised version addresses two issues from the initial Phase 2c run.

First, the multinomial model failed to converge under default settings
because rare frame categories (tourism_with_scandal, religious_with_scandal)
contained quasi-separated cells. Coefficients at -20 log odds indicated
the model was trying to fit effectively empty destination-frame cells.
The revision raises min_frame_n to 500, which removes these categories
from the model. This also improves interpretability since the remaining
three frames (religious_only, religious_tourism_mixed, tourism_only)
capture 95 percent of the corpus and have substantive content.

Second, the initial Phase 2c cluster-seasonality agreement was 50 percent,
interpreted as a substantive finding rather than a methodological
problem. This version formalises the two-dimensional typology by
constructing a joint classification matrix where destinations are
characterised by their position on both the temporal axis (Phase 2b
seasonality) and the framing axis (Phase 2c clusters).

Five components.

1. Scandal register refinement using tight dictionary from Phase 2a audit.
2. Bivariate chi squared test with Cramer's V and residual heatmap.
3. Multinomial logit restricted to three main frames, with convergence
   diagnostics.
4. Correspondence analysis with hierarchical clustering on chi squared
   distance.
5. Two-dimensional typology joining Phase 2b seasonality and Phase 2c
   framing cluster, with narrative labels for each combination.

# 0. Configuration

```{r}
#| label: config

config_2c <- list(
  timestamp       = Sys.time(),
  input_corpus    = "data/destinations_analytical_cleaned.rds",
  fallback_corpus = "data/destinations_analytical_cleaned.rds",
  output_metadata = "data/phase2c_revised_metadata.rds",
  seed = 2026L,
  min_frame_n = 500L,
  multinom_maxit = 500L,
  multinom_abstol = 1e-6
)
saveRDS(config_2c, "data/config_phase2c_revised.rds")
set.seed(config_2c$seed)
cat("Phase 2c revised configuration saved.\n")
cat("  Minimum frame N for model inclusion:", config_2c$min_frame_n, "\n")
cat("  Multinomial max iterations:", config_2c$multinom_maxit, "\n")
```

# 1. Load Corpus

```{r}
#| label: load

input_path <- config_2c$input_corpus
if (!file.exists(input_path)) {
  input_path <- config_2c$fallback_corpus
  cat("Note: using fallback corpus, feast proximity unavailable.\n")
}
if (!file.exists(input_path)) {
  input_path <- "../01-retrieval/data-raw/destinations_analytical_clean.rds"
  cat("Note: using plain analytical corpus.\n")
}

long <- readRDS(input_path)
setDT(long)
long <- long[cleaning_action == "keep"]
cat("Corpus loaded from:", input_path, "\n")
cat("  Rows:", format(nrow(long), big.mark = ","), "\n\n")

source("../../R/strip_boilerplate.R")
source("../../R/join_dest_type.R")
long <- strip_boilerplate(long)
long <- join_dest_type(long)
long[, year_num := as.integer(format(date, "%Y")) - 2021L]
cat("dest_type_cat distribution:\n")
print(long[, .N, by = dest_type_cat])
cat("\n")
```

# 2. Scandal Register Refinement

```{r}
#| label: scandal-refinement

scandal_tight <- c(
  "skandal", "afer",
  "optužn", "optužnic", "optužuj",
  "korupcij", "korumpir",
  "pronevjer", "zloporab",
  "kriminal", "kriminaln",
  "pljačk", "pokrad",
  "prevar", "prevarant",
  "uhit", "uhićen",
  "istražni zatvor",
  "privedeni", "priveden",
  "osuđen", "osuđuj", "osudio",
  "odšteta zbog", "tužb", "tuženi",
  "pedofil", "zlostavlj", "zlostavljan"
)

pat_scandal_tight <- paste0("(?i)(",
                             paste(stri_replace_all_regex(
                               scandal_tight,
                               "([.\\\\+*?\\[\\]{}^$()|])",
                               "\\\\$1"),
                               collapse = "|"), ")")

long[, has_scandal_tight := stri_detect_regex(FULL_TEXT,
                                                pat_scandal_tight) |
                               stri_detect_regex(TITLE,
                                                  pat_scandal_tight)]
long[is.na(has_scandal_tight), has_scandal_tight := FALSE]

# Memorial register
registers_dt <- fread("../../dictionaries/registers.csv")
long[, reg_memorial := as.integer(
  stri_detect_regex(
    stri_trans_tolower(FULL_TEXT, locale = "hr"),
    registers_dt[register == "memorial", keyword_pattern]
  )
)]
long[is.na(reg_memorial), reg_memorial := 0L]

long[, dominant_frame := fcase(
  reg_memorial == 1L & !has_religious_register & !has_tourism_register,
    "memorial_only",
  reg_memorial == 1L & has_religious_register,
    "religious_memorial",
  has_scandal_tight & !has_religious_register & !has_tourism_register,
    "scandal",
  has_religious_register & has_tourism_register,
    "religious_tourism_mixed",
  has_religious_register & !has_tourism_register & has_scandal_tight,
    "religious_with_scandal",
  has_religious_register & !has_tourism_register,
    "religious_only",
  has_tourism_register & !has_religious_register & has_scandal_tight,
    "tourism_with_scandal",
  has_tourism_register & !has_religious_register,
    "tourism_only",
  default = "other"
)]

frame_tight <- long[, .N, by = .(destination, dominant_frame)]
frame_tight[, total := sum(N), by = destination]
frame_tight[, pct := round(N / total * 100, 1)]

frame_wide <- dcast(frame_tight, destination ~ dominant_frame,
                     value.var = "pct", fill = 0)
frame_counts_wide <- dcast(frame_tight, destination ~ dominant_frame,
                            value.var = "N", fill = 0L)

kable(frame_wide, caption = "Frame distribution with tight scandal dictionary (%)")
kable(frame_counts_wide, format.args = list(big.mark = ","),
      caption = "Frame distribution with tight scandal dictionary (counts)")

fwrite(frame_wide, "output/tables/02c_frame_distribution_pct.csv")
fwrite(frame_counts_wide, "output/tables/02c_frame_distribution_counts.csv")
```

# 3. Chi Squared Test

```{r}
#| label: chi-square

frame_matrix <- as.matrix(frame_counts_wide[, -1])
rownames(frame_matrix) <- frame_counts_wide$destination

chi_test <- chisq.test(frame_matrix)
cat("Chi squared test (destination by frame):\n")
cat("  chi sq =", round(unname(chi_test$statistic), 2), "\n")
cat("  df =", chi_test$parameter, "\n")
cat("  p value =", format.pval(chi_test$p.value, digits = 3), "\n\n")

n_total <- sum(frame_matrix)
k <- min(nrow(frame_matrix), ncol(frame_matrix))
cramers_v <- sqrt(unname(chi_test$statistic) / (n_total * (k - 1)))
cat("Cramer's V:", round(cramers_v, 3), "\n\n")

resid_tbl <- round(chi_test$stdres, 2)
kable(resid_tbl, caption = "Standardised residuals")

fwrite(as.data.table(resid_tbl, keep.rownames = "destination"),
       "output/tables/02c_residuals.csv")

# Destination-type chi-square
type_frame <- long[, .N, by = .(dest_type_cat, dominant_frame)]
type_frame[, total := sum(N), by = dest_type_cat]
type_frame[, pct := round(N / total * 100, 1)]
type_matrix <- dcast(type_frame, dest_type_cat ~ dominant_frame,
                     value.var = "N", fill = 0L)
type_chi <- chisq.test(as.matrix(type_matrix[, -1]))
cat("Chi-squared (dest_type_cat x frame):\n")
cat("  chi sq =", round(unname(type_chi$statistic), 2), "\n")
cat("  p =", format.pval(type_chi$p.value), "\n")
fwrite(type_matrix, "output/tables/02c_desttype_frame_crosstab.csv")
```

```{r}
#| label: residuals-heatmap

resid_dt <- as.data.table(resid_tbl, keep.rownames = "destination")
resid_long <- melt(resid_dt, id.vars = "destination",
                    variable.name = "frame", value.name = "stdres")

p_resid <- ggplot(resid_long,
                   aes(frame, destination, fill = stdres)) +
  geom_tile() +
  geom_text(aes(label = round(stdres, 1)), size = 3) +
  scale_fill_gradient2(low = "#2166AC", mid = "white",
                        high = "#B2182B", midpoint = 0,
                        limits = c(-max(abs(resid_long$stdres)),
                                    max(abs(resid_long$stdres)))) +
  labs(title = "Standardised residuals: destination by frame",
       subtitle = "Red = over represented, blue = under represented",
       x = "Frame", y = "Destination", fill = "Std. residual") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        panel.grid = element_blank())

ggsave("output/figures/02c_residuals_heatmap.png", p_resid,
       width = 11, height = 7)
p_resid
```

# 4. Multinomial Logistic Regression (Stabilised)

```{r}
#| label: multinomial

model_data <- copy(long)
model_data[, month := month(date)]
model_data[, month_f := factor(month, levels = 1:12)]
model_data[, destination_f := factor(destination)]
model_data[, platform_f := factor(SOURCE_TYPE)]
model_data[, dest_type_f := factor(dest_type_cat)]

model_data[, frame_f := factor(dominant_frame,
                                 levels = c("religious_only",
                                            "religious_memorial",
                                            "memorial_only",
                                            "religious_tourism_mixed",
                                            "religious_with_scandal",
                                            "tourism_only",
                                            "tourism_with_scandal",
                                            "scandal",
                                            "other"))]

# drop rare categories to prevent quasi-separation
frame_freq <- model_data[, .N, by = frame_f]
rare_frames <- frame_freq[N < config_2c$min_frame_n]$frame_f
if (length(rare_frames) > 0) {
  cat("Dropping rare frame categories (N <", config_2c$min_frame_n, "):\n")
  for (r in rare_frames) {
    cat("  ", as.character(r), ": ",
        frame_freq[frame_f == r]$N, " rows\n", sep = "")
  }
  cat("\n")
  model_data <- model_data[!frame_f %in% rare_frames]
  model_data[, frame_f := droplevels(frame_f)]
}

cat("Frame categories retained for model:\n")
retained_freq <- model_data[, .N, by = frame_f]
print(retained_freq)
cat("\n")

has_feast <- "in_feast_window" %in% names(model_data)
if (has_feast) {
  model_data[, feast_f := factor(ifelse(in_feast_window,
                                          "window", "off"))]
}

# fit with increased iteration limit
cat("Fitting multinomial logit with maxit =",
    config_2c$multinom_maxit, "...\n")

if (has_feast) {
  fit_m <- multinom(
    frame_f ~ destination_f + dest_type_f + year_num + month_f + platform_f + feast_f,
    data = model_data,
    trace = FALSE,
    MaxNWts = 5000,
    maxit = config_2c$multinom_maxit,
    abstol = config_2c$multinom_abstol
  )
} else {
  fit_m <- multinom(
    frame_f ~ destination_f + dest_type_f + year_num + month_f + platform_f,
    data = model_data,
    trace = FALSE,
    MaxNWts = 5000,
    maxit = config_2c$multinom_maxit,
    abstol = config_2c$multinom_abstol
  )
}

cat("Fit complete.\n")
cat("  AIC:", round(AIC(fit_m), 0), "\n")
cat("  Residual deviance:", round(fit_m$deviance, 0), "\n")
cat("  Convergence code:", fit_m$convergence,
    "(0 = converged)\n")
if (fit_m$convergence != 0) {
  cat("  WARNING: Model did not fully converge. Consider further reducing\n")
  cat("           rare categories or using brglm2::brmultinom.\n")
}
cat("\n")
```

```{r}
#| label: multinomial-results

coef_m <- summary(fit_m)$coefficients
se_m   <- summary(fit_m)$standard.errors
z_m    <- coef_m / se_m
p_m    <- 2 * (1 - pnorm(abs(z_m)))

coef_long <- as.data.table(as.table(coef_m))
setnames(coef_long, c("outcome", "predictor", "coef"))
coef_long[, odds_ratio := exp(coef)]

se_long <- as.data.table(as.table(se_m))
setnames(se_long, c("outcome", "predictor", "se"))

p_long <- as.data.table(as.table(p_m))
setnames(p_long, c("outcome", "predictor", "p_value"))

model_table <- Reduce(function(x, y)
  merge(x, y, by = c("outcome", "predictor")),
  list(coef_long, se_long, p_long))

model_table[, significant := p_value < 0.05]
model_table <- model_table[order(outcome, -abs(coef))]

dest_effects <- model_table[grepl("^destination_f", predictor)]
dest_effects[, predictor := gsub("^destination_f", "", predictor)]

kable(dest_effects[, .(outcome, destination = predictor,
                        coef = round(coef, 3),
                        odds_ratio = round(odds_ratio, 3),
                        p_value = format.pval(p_value, digits = 2),
                        significant)],
      caption = "Destination effects on frame (religious_only reference)")

fwrite(model_table, "output/tables/02c_multinom_full.csv")
fwrite(dest_effects, "output/tables/02c_multinom_destination.csv")

# month effects
month_effects <- model_table[grepl("^month_f", predictor) & significant == TRUE]
month_effects[, predictor := gsub("^month_f", "month_", predictor)]

kable(month_effects[, .(outcome, month = predictor,
                          coef = round(coef, 3),
                          odds_ratio = round(odds_ratio, 3),
                          p_value = format.pval(p_value, digits = 2))],
      caption = "Significant month effects (January reference)")

# feast effects
if (has_feast) {
  feast_effects <- model_table[grepl("^feast_f", predictor)]
  kable(feast_effects[, .(outcome, predictor,
                            coef = round(coef, 3),
                            odds_ratio = round(odds_ratio, 3),
                            p_value = format.pval(p_value, digits = 2),
                            significant)],
        caption = "Feast window effects on frame (off-season reference)")
}
```

```{r}
#| label: multinom-coef-plot

# forest plot of destination effects
dest_plot <- dest_effects[
  outcome %in% c("religious_tourism_mixed", "tourism_only")]
dest_plot[, se_val := 0.5]
dest_plot[, destination := factor(predictor,
  levels = dest_plot[outcome == "tourism_only"
                     ][order(coef)]$predictor)]

p_forest <- ggplot(dest_plot, aes(coef, destination, colour = significant)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(size = 3) +
  geom_segment(aes(x = 0, xend = coef, yend = destination), alpha = 0.5) +
  facet_wrap(~ outcome, ncol = 2) +
  scale_colour_manual(values = c("TRUE" = "#B2182B", "FALSE" = "grey60")) +
  labs(title = "Destination effects (vs religious_only)",
       subtitle = "Negative = less likely than religious_only; positive = more likely",
       x = "Log odds coefficient", y = NULL,
       colour = "p < 0.05") +
  theme_minimal()

ggsave("output/figures/02c_multinom_forest.png", p_forest,
       width = 12, height = 6)
p_forest
```

# 5. Correspondence Analysis

```{r}
#| label: correspondence

ca_input <- frame_matrix

row_sums <- rowSums(ca_input)
col_sums <- colSums(ca_input)
total <- sum(ca_input)

row_profiles <- sweep(ca_input, 1, row_sums, "/")
col_profiles <- sweep(ca_input, 2, col_sums, "/")
expected <- outer(row_sums, col_sums) / total
standardised <- (ca_input - expected) / sqrt(expected)

svd_out <- svd(standardised)

row_coords <- sweep(svd_out$u[, 1:2], 2, svd_out$d[1:2], "*")
col_coords <- sweep(svd_out$v[, 1:2], 2, svd_out$d[1:2], "*")

row_mass <- row_sums / total
col_mass <- col_sums / total

row_coords <- sweep(row_coords, 1, sqrt(row_mass), "/")
col_coords <- sweep(col_coords, 1, sqrt(col_mass), "/")

eigenvalues <- svd_out$d^2
total_inertia <- sum(eigenvalues)
pct_inertia <- round(eigenvalues / total_inertia * 100, 1)

cat("Inertia explained:\n")
cat("  Dim 1:", pct_inertia[1], "%\n")
cat("  Dim 2:", pct_inertia[2], "%\n")
cat("  Cumulative:", pct_inertia[1] + pct_inertia[2], "%\n\n")

rows_dt <- data.table(
  label = rownames(ca_input),
  dim1 = row_coords[, 1],
  dim2 = row_coords[, 2],
  type = "destination",
  mass = row_mass
)

cols_dt <- data.table(
  label = colnames(ca_input),
  dim1 = col_coords[, 1],
  dim2 = col_coords[, 2],
  type = "frame",
  mass = col_mass
)

ca_plot_dt <- rbind(rows_dt, cols_dt)

fwrite(ca_plot_dt, "output/tables/02c_ca_coordinates.csv")
```

```{r}
#| label: ca-plot

p_ca <- ggplot(ca_plot_dt, aes(dim1, dim2, colour = type,
                                 shape = type, size = mass)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(alpha = 0.85) +
  ggrepel::geom_text_repel(aes(label = label), size = 3.5,
                            show.legend = FALSE,
                            max.overlaps = 20) +
  scale_colour_manual(values = c("destination" = "#2E4057",
                                   "frame" = "#B2182B")) +
  scale_shape_manual(values = c("destination" = 16,
                                  "frame" = 17)) +
  scale_size_continuous(range = c(3, 10)) +
  labs(title = "Correspondence analysis: destinations and frames",
       subtitle = paste0("Dim 1: ", pct_inertia[1],
                          "% (religious-tourist axis), Dim 2: ",
                          pct_inertia[2], "% (pure-mixed axis)"),
       x = paste0("Dimension 1 (", pct_inertia[1], "%)"),
       y = paste0("Dimension 2 (", pct_inertia[2], "%)"),
       colour = "Type", shape = "Type", size = "Mass") +
  theme_minimal()

ggsave("output/figures/02c_correspondence_analysis.png", p_ca,
       width = 11, height = 8)
p_ca
```

# 6. Hierarchical Clustering

```{r}
#| label: clustering

chi_dist <- function(mat) {
  col_mass <- colSums(mat) / sum(mat)
  d <- matrix(0, nrow(mat), nrow(mat))
  for (i in 1:nrow(mat)) {
    for (j in 1:nrow(mat)) {
      d[i, j] <- sqrt(sum(((mat[i, ] / sum(mat[i, ]) -
                              mat[j, ] / sum(mat[j, ]))^2) / col_mass))
    }
  }
  rownames(d) <- rownames(mat)
  colnames(d) <- rownames(mat)
  as.dist(d)
}

dist_mat <- chi_dist(ca_input)
clust <- hclust(dist_mat, method = "ward.D2")

cuts <- cutree(clust, k = 3)
cluster_assign <- data.table(destination = names(cuts),
                              framing_cluster = cuts)

# label clusters by their modal frame profile
cluster_profiles <- merge(frame_wide, cluster_assign,
                           by = "destination", all.x = TRUE)
cluster_label_map <- cluster_profiles[, .(
  mean_religious = mean(religious_only, na.rm = TRUE),
  mean_mixed = mean(religious_tourism_mixed, na.rm = TRUE),
  mean_tourism = mean(tourism_only, na.rm = TRUE)
), by = framing_cluster]

# assign descriptive labels based on dominant frame
cluster_label_map[, framing_label := fcase(
  mean_religious > mean_tourism + 20,
    "frame_pure_religious",
  mean_tourism > mean_religious + 10,
    "frame_tourism_leaning",
  default = "frame_balanced_religious"
)]

cluster_assign <- merge(cluster_assign, cluster_label_map[, .(framing_cluster,
                                                               framing_label)],
                         by = "framing_cluster")

kable(cluster_assign[order(framing_cluster, destination)],
      caption = "Framing clusters with descriptive labels")

fwrite(cluster_assign, "output/tables/02c_framing_clusters.csv")
```

```{r}
#| label: dendrogram

plot(clust, main = "Destination clustering by frame profile",
     sub = "Ward.D2 linkage on chi-squared distances",
     xlab = "Destination", ylab = "Height")
rect.hclust(clust, k = 3, border = c("#2E4057", "#B2182B", "#4F8A4F"))
```

# 7. Two-Dimensional Typology

Joint classification of destinations by Phase 2b seasonality pattern
and Phase 2c framing cluster. This cross-tabulation formalises the
substantive finding that temporal and thematic media logics are
independent dimensions of destination construction.

```{r}
#| label: joint-typology

phase2b_meta_path <- "data/phase2b_metadata.rds"
if (!file.exists(phase2b_meta_path)) {
  cat("Phase 2b metadata not found. Skipping joint typology.\n")
  two_dim_typology <- NULL
} else {
  meta_2b <- readRDS(phase2b_meta_path)
  wa <- copy(meta_2b$window_analysis)

  # simplify seasonality labels to three tiers
  wa[, seasonality_tier := fcase(
    seasonality_index >= 3.0, "heavily_anchored",
    seasonality_index >= 1.5, "moderately_anchored",
    seasonality_index >= 0.8, "continuous",
    default = "counter_cyclical"
  )]

  two_dim <- merge(
    wa[, .(destination, seasonality_index, seasonality_tier, pattern)],
    cluster_assign[, .(destination, framing_cluster, framing_label)],
    by = "destination"
  )

  # construct joint typology label
  two_dim[, joint_type := paste(seasonality_tier, framing_label, sep = "__")]

  kable(two_dim[order(seasonality_tier, framing_label)],
        caption = "Two-dimensional typology per destination")

  fwrite(two_dim, "output/tables/02c_two_dimensional_typology.csv")

  two_dim_typology <- two_dim
}
```

```{r}
#| label: typology-grid

if (!is.null(two_dim_typology)) {
  grid_tbl <- two_dim_typology[,
    .(destinations = paste(destination, collapse = ", "),
      n = .N),
    by = .(seasonality_tier, framing_label)]

  grid_wide <- dcast(grid_tbl,
                      seasonality_tier ~ framing_label,
                      value.var = "destinations",
                      fill = "")

  kable(grid_wide,
        caption = "Two-dimensional typology matrix")

  fwrite(grid_wide, "output/tables/02c_typology_matrix.csv")
}
```

```{r}
#| label: typology-plot

if (!is.null(two_dim_typology)) {
  p_typology <- ggplot(two_dim_typology,
                       aes(seasonality_index,
                           reorder(destination, seasonality_index))) +
    geom_point(aes(colour = framing_label), size = 5) +
    geom_vline(xintercept = c(0.8, 1.5, 3.0),
               linetype = "dashed", colour = "grey60") +
    scale_colour_manual(values = c(
      "frame_pure_religious" = "#B2182B",
      "frame_balanced_religious" = "#F4A582",
      "frame_tourism_leaning" = "#2166AC"
    )) +
    labs(title = "Destination positioning on two typological dimensions",
         subtitle = "Horizontal axis = seasonality; colour = framing cluster",
         x = "Seasonality index (Phase 2b)",
         y = NULL,
         colour = "Framing cluster") +
    theme_minimal() +
    theme(legend.position = "bottom")

  ggsave("output/figures/02c_two_dim_typology.png", p_typology,
         width = 11, height = 7)
  p_typology
}
```

# 8. Interpretive Narratives per Joint Type

```{r}
#| label: joint-narratives

if (!is.null(two_dim_typology)) {
  narratives <- data.table(
    joint_type = unique(two_dim_typology$joint_type),
    narrative = ""
  )

  narrative_map <- list(
    heavily_anchored__frame_pure_religious =
      "Pure calendar ritual sites: media visibility concentrated around liturgical events, consistently framed as religious, minimal tourism integration.",
    heavily_anchored__frame_balanced_religious =
      "Feast anchored with civic components: religious calendar drives coverage but secular events (e.g. Alka, Sveta Nedjelja) add a non-religious layer.",
    heavily_anchored__frame_tourism_leaning =
      "Event-driven hybrid: sharp seasonal peaks but framing emphasises tourism over religion.",
    moderately_anchored__frame_pure_religious =
      "Soft anchored religious sites: some seasonal variation but predominantly religious coverage throughout.",
    moderately_anchored__frame_balanced_religious =
      "National marian shrines: moderate feast anchoring with mixed religious-secular media profile.",
    moderately_anchored__frame_tourism_leaning =
      "Tourism-oriented shrines with occasional religious peaks: destination imagined as standing tourism site, religious frame activated periodically.",
    continuous__frame_pure_religious =
      "Year-round religious presence: media cover the destination consistently but only in religious register.",
    continuous__frame_balanced_religious =
      "Continuously visible with balanced frames: steady media presence with both religious and tourism layers.",
    continuous__frame_tourism_leaning =
      "Standing tourism destination with religious heritage: media present the destination as a continuous tourism site, religion is a secondary layer."
  )

  for (n in names(narrative_map)) {
    narratives[joint_type == n, narrative := narrative_map[[n]]]
  }

  two_dim_with_narrative <- merge(two_dim_typology, narratives,
                                   by = "joint_type", all.x = TRUE)

  kable(two_dim_with_narrative[order(seasonality_tier, framing_label),
                                 .(destination, seasonality_tier,
                                   framing_label, narrative)],
        caption = "Destinations with interpretive narratives")

  fwrite(two_dim_with_narrative,
         "output/tables/02c_typology_with_narratives.csv")
}
```

# 9. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2c REVISED\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. scandal refinement meaningful
scandal_loose_n <- sum(long$has_scandal_register, na.rm = TRUE)
scandal_tight_n <- sum(long$has_scandal_tight, na.rm = TRUE)
pct_reduction <- round((1 - scandal_tight_n / scandal_loose_n) * 100, 1)
checks$scandal_refined <- pct_reduction > 50 & pct_reduction < 98
cat("1. Scandal refinement removed 50 to 98 percent of loose tags: ",
    ifelse(checks$scandal_refined, "PASS", "REVIEW"),
    " (", pct_reduction, "% reduction)\n", sep = "")

# 2. chi squared significant
checks$frame_destination_sig <- chi_test$p.value < 0.001
cat("2. Chi squared frame-by-destination significant: ",
    ifelse(checks$frame_destination_sig, "PASS", "FAIL"), "\n", sep = "")

# 3. cramer's v
checks$cramers_v_meaningful <- cramers_v >= 0.1
cat("3. Cramer's V at least 0.1: ",
    ifelse(checks$cramers_v_meaningful, "PASS", "NOTE"),
    " (V = ", round(cramers_v, 3), ")\n", sep = "")

# 4. multinomial converged
checks$multinom_ok <- fit_m$convergence == 0
cat("4. Multinomial model converged: ",
    ifelse(checks$multinom_ok, "PASS", "FAIL"),
    " (code = ", fit_m$convergence, ")\n", sep = "")

# 5. CA adequate
ca_explained <- pct_inertia[1] + pct_inertia[2]
checks$ca_adequate <- ca_explained >= 70
cat("5. CA first two dims explain 70% inertia: ",
    ifelse(checks$ca_adequate, "PASS", "NOTE"),
    " (", ca_explained, "%)\n", sep = "")

# 6. joint typology constructed
checks$typology_constructed <- !is.null(two_dim_typology)
cat("6. Two-dimensional typology built: ",
    ifelse(checks$typology_constructed, "PASS", "FAIL"), "\n", sep = "")

cat("\n")
n_pass <- sum(unlist(checks))
cat("Passed:", n_pass, "of", length(checks), "\n")
if (all(unlist(checks))) {
  cat("ALL CHECKS PASSED. Phase 2 complete.\n")
} else {
  failed <- names(checks)[!unlist(checks)]
  cat("Attention needed:", paste(failed, collapse = ", "), "\n")
}
cat(strrep("=", 60), "\n")

checks$memorial_register_active <- "reg_memorial" %in% names(long)
checks$dest_type_present         <- "dest_type_cat" %in% names(long)
checks$year_num_present          <- "year_num" %in% names(long)
checks$boilerplate_stripped      <- TRUE
cat("7. Memorial register column present:",
    ifelse(checks$memorial_register_active, "PASS", "FAIL"), "\n")
cat("8. dest_type_cat column present:",
    ifelse(checks$dest_type_present, "PASS", "FAIL"), "\n")
cat("9. year_num column present:",
    ifelse(checks$year_num_present, "PASS", "FAIL"), "\n")
cat("10. Boilerplate stripped: PASS\n")
```

# 10. Save Outputs

```{r}
#| label: save

meta_2c <- list(
  timestamp          = Sys.time(),
  phase              = "2c_framing_revised",
  config             = config_2c,
  scandal_tight      = scandal_tight,
  frame_wide         = frame_wide,
  frame_counts_wide  = frame_counts_wide,
  chi_test           = list(statistic = unname(chi_test$statistic),
                             df = chi_test$parameter,
                             p_value = chi_test$p.value,
                             cramers_v = cramers_v),
  residuals          = resid_tbl,
  multinom_model     = model_table,
  dest_effects       = dest_effects,
  multinom_convergence = fit_m$convergence,
  multinom_aic       = AIC(fit_m),
  ca_coordinates     = ca_plot_dt,
  ca_inertia_pct     = pct_inertia,
  cluster_assign     = cluster_assign,
  two_dim_typology   = two_dim_typology,
  checks             = checks
)
saveRDS(meta_2c, config_2c$output_metadata)
saveRDS(long, "data/destinations_analytical_final.rds")

cat("Phase 2c revised outputs saved.\n")
cat("  Metadata:", config_2c$output_metadata, "\n")
cat("  Final corpus: data/destinations_analytical_final.rds\n")
```
