---
title: "Phase 2d: Keyness Analysis"
subtitle: "Textual profiles of destinations via log-likelihood keyness"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(stringi)
library(quanteda)
library(quanteda.textstats)
library(ggplot2)
library(knitr)

for (d in c("data", "output/tables", "output/figures")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 2d produces textual profiles of destinations. For each destination,
the log likelihood keyness test identifies tokens that are statistically
over represented in articles about that destination compared to the rest
of the corpus. Output is a ranked list of characteristic vocabulary per
destination, forming the inductive counterpart to the deductive framing
categories from Phase 2c.

Keyness fills a specific role in the analytical pipeline. Framing
analysis operates on predefined dictionary categories that capture
thematic structure at the level of registers. Keyness operates on the
full text corpus without prior assumptions, revealing destination
specific vocabulary that may or may not align with dictionary
categories. The two approaches triangulate, keyness empirically
validates the registers where alignment holds and identifies residual
lexical variation where it does not.

The keyness statistic used is the log likelihood ratio G² (Dunning 1993),
which is more robust than chi squared for low frequency tokens. Values
above 10 are conventionally interpreted as statistically meaningful, and
values above 100 as strongly distinctive.

# 0. Configuration

```{r}
#| label: config

config_2d <- list(
  timestamp           = Sys.time(),
  input_corpus        = "data/destinations_analytical_cleaned.rds",
  fallback_corpus     = "data/destinations_analytical_cleaned.rds",
  output_metadata     = "data/phase2d_metadata.rds",
  top_n_keywords      = 25,
  min_token_length    = 3,
  min_doc_frequency   = 10,
  min_target_docs     = 50,
  seed                = 2026L
)
set.seed(config_2d$seed)
saveRDS(config_2d, "data/config_phase2d.rds")
cat("Phase 2d configuration saved.\n")
```

# 1. Load Corpus

```{r}
#| label: load

input_path <- config_2d$input_corpus
if (!file.exists(input_path)) input_path <- config_2d$fallback_corpus

long <- readRDS(input_path)
setDT(long)
long <- long[cleaning_action == "keep"]

cat("Corpus loaded from:", input_path, "\n")
cat("  Total pairs:", format(nrow(long), big.mark = ","), "\n")
cat("  Unique articles:",
    format(uniqueN(long$article_id), big.mark = ","), "\n\n")
```

# 2. Article Level Text Base

Keyness must operate on unique articles, not on pairs. An article that
mentions two destinations would otherwise inflate both destinations'
token counts artificially. The construction below takes each article's
text once and maintains a separate article-to-destination mapping.

```{r}
#| label: article-base

# keep only articles with usable text
long_valid <- long[!is.na(FULL_TEXT) & nchar(FULL_TEXT) > 100]

articles <- unique(long_valid[, .(article_id, FULL_TEXT)],
                    by = "article_id")

# destination membership per article (one article can belong to multiple)
art_dest <- unique(long_valid[, .(article_id, destination)])

cat("Article base constructed.\n")
cat("  Unique articles with text:",
    format(nrow(articles), big.mark = ","), "\n")
cat("  Article-destination mappings:",
    format(nrow(art_dest), big.mark = ","), "\n\n")

dest_article_counts <- art_dest[, .N, by = destination][order(-N)]
kable(dest_article_counts, format.args = list(big.mark = ","),
      caption = "Unique articles per destination")
```

# 3. Tokenisation and DFM

```{r}
#| label: tokenisation

croatian_stops <- c(
  "biti", "jesam", "jesi", "jest", "jesmo", "jeste", "jesu",
  "sam", "si", "je", "smo", "ste", "su",
  "bio", "bila", "bilo", "bili", "bile",
  "budem", "budeš", "bude", "budemo", "budete", "budu",
  "bih", "bi", "bismo", "biste",
  "nisam", "nisi", "nije", "nismo", "niste", "nisu",
  "ga", "mu", "joj", "ih", "im", "mi", "ti", "me",
  "te", "mene", "tebe", "on", "ona", "ono", "oni", "one",
  "njega", "njemu", "njoj", "njima", "nama", "vama", "nas", "vas",
  "koji", "koja", "koje", "kojim", "kojima", "kojem", "kojeg",
  "kako", "kada", "gdje", "zašto", "dok", "dokle", "otkad",
  "tako", "ovako", "onako", "sada", "tada", "onda", "još",
  "samo", "neki", "neka", "neko", "nekad", "nešto", "netko",
  "ako", "jer", "što", "čemu", "čega", "čime", "čim",
  "preko", "iz", "od", "do", "za", "po",
  "prije", "poslije", "tijekom", "pri", "bez", "nakon", "prema",
  "među", "između", "iznad", "ispod", "pored", "uz",
  "ovaj", "ova", "ovo", "ovi", "ove", "ovim", "ovu",
  "taj", "ta", "to", "te", "tim", "tu", "ovoga", "toga",
  "onaj", "ono", "onim",
  "njegov", "njezin", "njihov", "njegova", "njezina",
  "sve", "svi", "svih", "svoj", "svoja", "svoje", "svaki",
  "god", "godine", "godina", "danas", "jučer", "sutra",
  "više", "manje", "vrlo", "već", "može", "mora", "treba",
  "kaže", "rekao", "rekla", "kazao", "kazala", "navodi", "naveo",
  "imati", "ima", "nema", "imao", "imala", "imali",
  "prva", "prvi", "prvo", "druga", "drugi", "drugo",
  "puno", "malo", "posebno", "osobito",
  "radi", "zbog", "ili", "ali", "ili", "jer",
  "http", "https", "www", "com", "hr", "rtl", "html"
)

qc <- corpus(articles, text_field = "FULL_TEXT", docid_field = "article_id")

toks <- tokens(qc,
                what = "word",
                remove_punct = TRUE,
                remove_numbers = TRUE,
                remove_symbols = TRUE,
                remove_url = TRUE,
                split_hyphens = TRUE) |>
  tokens_tolower() |>
  tokens_select(min_nchar = config_2d$min_token_length) |>
  tokens_remove(croatian_stops)

dfm_d <- dfm(toks) |>
  dfm_trim(min_docfreq = config_2d$min_doc_frequency)

cat("DFM constructed.\n")
cat("  Documents:", format(ndoc(dfm_d), big.mark = ","), "\n")
cat("  Vocabulary:", format(nfeat(dfm_d), big.mark = ","), "\n\n")
```

# 4. Per Destination Keyness

```{r}
#| label: keyness-loop

destinations <- unique(long$destination)
keyness_list <- list()

for (d in destinations) {
  target_articles <- art_dest[destination == d]$article_id
  is_target <- docnames(dfm_d) %in% target_articles

  n_target <- sum(is_target)
  if (n_target < config_2d$min_target_docs) {
    cat("Skipping", d, "only", n_target, "target articles.\n")
    next
  }

  key_result <- textstat_keyness(dfm_d, target = is_target,
                                   measure = "lr")

  key_top <- head(key_result, config_2d$top_n_keywords)
  key_top_dt <- as.data.table(key_top)
  key_top_dt[, destination := d]
  key_top_dt[, rank := .I]
  keyness_list[[d]] <- key_top_dt

  cat(sprintf("  %-18s | %5s target articles | top G2 = %s\n",
              d, format(n_target, big.mark = ","),
              round(max(key_top_dt$G2), 1)))
}

all_keyness <- rbindlist(keyness_list, fill = TRUE)
fwrite(all_keyness, "output/tables/02d_keyness_all.csv")
```

# 5. Per Destination Tables

```{r}
#| label: display-tables
#| results: asis

for (d in destinations) {
  if (!d %in% all_keyness$destination) next
  cat("\n\n## ", d, "\n\n", sep = "")
  show <- all_keyness[destination == d,
                        .(rank, feature,
                          G2 = round(G2, 1),
                          p_value = format.pval(p, digits = 2),
                          n_target, n_reference)]
  print(kable(show, caption = paste("Top keywords for", d)))
}
```

# 6. Cross Destination Overlap

Most keywords should be specific to one destination. High overlap across
destinations would suggest the distinctive vocabulary is thin.

```{r}
#| label: overlap

shared_keywords <- all_keyness[, .N, by = feature][order(-N)]
setnames(shared_keywords, c("feature", "n_destinations"))

kable(shared_keywords[n_destinations >= 3],
      caption = "Keywords appearing in three or more destinations")

pct_unique <- round(sum(shared_keywords$n_destinations == 1) /
                      nrow(shared_keywords) * 100, 1)
cat("\nPercent of keywords appearing in only one destination: ",
    pct_unique, "%\n", sep = "")

fwrite(shared_keywords, "output/tables/02d_shared_keywords.csv")
```

# 7. Keyness Heatmap

```{r}
#| label: heatmap

top_words_union <- unique(all_keyness[rank <= 10]$feature)
heatmap_data <- all_keyness[feature %in% top_words_union]

# order features by destination they belong to
feature_order <- heatmap_data[, .(first_dest = destination[which.min(rank)],
                                    min_rank = min(rank)),
                               by = feature][order(first_dest, min_rank)]
heatmap_data[, feature := factor(feature,
                                   levels = feature_order$feature)]

p_heat <- ggplot(heatmap_data,
                  aes(destination, feature, fill = G2)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "#B2182B") +
  labs(title = "Keyness heatmap",
       subtitle = "Top 10 keywords per destination, G2 log-likelihood",
       x = NULL, y = NULL, fill = "G2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        axis.text.y = element_text(size = 7),
        panel.grid = element_blank())

ggsave("output/figures/02d_keyness_heatmap.png", p_heat,
       width = 11, height = 18)
p_heat
```

# 8. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2d\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. all destinations profiled
checks$all_destinations_profiled <- length(keyness_list) == length(destinations)
cat("1. All destinations profiled: ",
    ifelse(checks$all_destinations_profiled, "PASS", "NOTE"),
    " (", length(keyness_list), " of ", length(destinations), ")\n", sep = "")

# 2. meaningful G2 values in top 5 for each destination
meaningful <- all_keyness[rank <= 5][, all(G2 > 10), by = destination]
checks$meaningful_g2 <- all(meaningful$V1)
cat("2. All destinations have top 5 keywords with G2 > 10: ",
    ifelse(checks$meaningful_g2, "PASS", "REVIEW"), "\n", sep = "")

# 3. sufficient unique vocabulary
checks$unique_keywords <- pct_unique >= 60
cat("3. At least 60% of keywords are destination specific: ",
    ifelse(checks$unique_keywords, "PASS", "NOTE"),
    " (", pct_unique, "%)\n", sep = "")

# 4. keyness distribution spans reasonable range
g2_range <- diff(range(all_keyness$G2))
checks$g2_range <- g2_range > 100
cat("4. G2 range spans > 100 across corpus: ",
    ifelse(checks$g2_range, "PASS", "NOTE"),
    " (range = ", round(g2_range, 1), ")\n", sep = "")

cat("\n")
cat("Passed:", sum(unlist(checks)), "of", length(checks), "\n")
cat(strrep("=", 60), "\n")
```

# 9. Save Outputs

```{r}
#| label: save

meta_2d <- list(
  timestamp          = Sys.time(),
  phase              = "2d_keyness",
  config             = config_2d,
  all_keyness        = all_keyness,
  shared_keywords    = shared_keywords,
  pct_unique         = pct_unique,
  checks             = checks,
  n_articles_used    = ndoc(dfm_d),
  n_features         = nfeat(dfm_d)
)
saveRDS(meta_2d, config_2d$output_metadata)
cat("Phase 2d saved to", config_2d$output_metadata, "\n")
```

# 10. Interpretive Summary

The keyness output feeds directly into the Descriptive Analysis section
of the results, specifically into the *Textual Profiles of Destinations*
subsection. Three analytical moves are now possible.

First, each destination has an empirical vocabulary signature that can
be described narratively. Expected patterns include Ludbreg anchored in
eucharistic vocabulary, Vepric in Lourdes and Makarska imagery, Sinj in
Alka and equestrian vocabulary, Marija Bistrica in national pilgrimage
language, Nin in medieval Croatian statehood vocabulary.

Second, the keyness output validates or challenges the framing
categories. Where religious keywords dominate, the dictionary tag
religious_only is empirically grounded. Where tourism or political
keywords dominate, the dictionary tag tourism_only is equally grounded.
Destinations whose top keywords mix both categories validate the
religious_tourism_mixed frame.

Third, cross destination overlap reveals the intertextual structure of
the corpus. Keywords shared across three or more destinations point to
generic religious or tourism vocabulary. Keywords specific to one
destination point to the distinctive identity each shrine occupies in
the media ecosystem.
