---
title: "Phase 2e: Per-Platform Framing Decomposition"
subtitle: "Is destination framing platform-invariant or platform-mediated"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(ggplot2)
library(knitr)
library(patchwork)

for (d in c("data", "output/tables", "output/figures")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 2e tests whether the framing patterns identified in Phase 2c are
stable across platforms or mediated by platform specific editorial
logics. The analytical corpus is dominated by web portals (about 80
percent of pairs), but Facebook, Twitter and YouTube together carry a
non trivial share. If the framing profile of a destination differs
substantially between platforms, the Phase 2c aggregate analysis hides
platform specific signal and the conclusions need platform qualifiers.
If the framing profile is stable across platforms, destination identity
dominates platform identity and the aggregate analysis holds.

Two analytical components.

First, per platform framing distributions per destination, with chi
squared tests computed within each platform. This establishes whether
the destination by frame relationship remains statistically significant
within each platform separately.

Second, platform shift analysis per destination, computing total
variation distance between platform profiles. Destinations with low
shift have stable framing across platforms, destinations with high
shift have platform mediated framing.

# 0. Configuration

```{r}
#| label: config

config_2e <- list(
  timestamp          = Sys.time(),
  input_corpus       = "data/destinations_analytical_cleaned.rds",
  output_metadata    = "data/phase2e_metadata.rds",
  min_platform_pairs = 500L,
  min_dest_platform_pairs = 50L,
  seed = 2026L
)
set.seed(config_2e$seed)
saveRDS(config_2e, "data/config_phase2e.rds")
cat("Phase 2e configuration saved.\n")
```

# 1. Load Corpus

```{r}
#| label: load

long <- readRDS(config_2e$input_corpus)
setDT(long)
long <- long[cleaning_action == "keep"]

cat("Corpus loaded.\n")
cat("  Total pairs:", format(nrow(long), big.mark = ","), "\n\n")

if (!"dominant_frame" %in% names(long)) {
  stop("dominant_frame column missing. Run Phase 2c first.")
}
```

# 2. Platform Inventory

```{r}
#| label: platform-inventory

platform_counts <- long[, .N, by = SOURCE_TYPE][order(-N)]
platform_counts[, pct := round(N / sum(N) * 100, 1)]
kable(platform_counts, format.args = list(big.mark = ","),
      caption = "Pairs per platform")

main_platforms <- platform_counts[N >= config_2e$min_platform_pairs
                                    ]$SOURCE_TYPE
cat("\nPlatforms retained for analysis (>= ",
    config_2e$min_platform_pairs, " pairs):\n  ",
    paste(main_platforms, collapse = ", "), "\n", sep = "")

long_main <- long[SOURCE_TYPE %in% main_platforms]
cat("\nCorpus restricted to main platforms: ",
    format(nrow(long_main), big.mark = ","), " pairs\n", sep = "")
```

# 3. Frame Distribution Per Platform

```{r}
#| label: per-platform-dist

# focus on the three main frames used in Phase 2c multinomial
main_frames <- c("religious_only", "religious_tourism_mixed",
                  "tourism_only")

long_main_frames <- long_main[dominant_frame %in% main_frames]

platform_frame_tbl <- long_main_frames[,
  .N, by = .(SOURCE_TYPE, destination, dominant_frame)]

# total per platform-destination
platform_frame_tbl[, total := sum(N),
                    by = .(SOURCE_TYPE, destination)]
platform_frame_tbl[, pct := round(N / total * 100, 1)]

# display one table per platform
for (p in main_platforms) {
  sub_wide <- dcast(platform_frame_tbl[SOURCE_TYPE == p],
                     destination ~ dominant_frame,
                     value.var = "pct", fill = 0)
  cat("\n\n## Platform", p, "\n\n")
  print(kable(sub_wide,
              caption = paste("Frame distribution (%) on", p)))
}

fwrite(platform_frame_tbl, "output/tables/02e_per_platform_frames.csv")
```

# 4. Chi Squared Test Per Platform

```{r}
#| label: per-platform-chi

chi_by_platform <- list()

for (p in main_platforms) {
  sub <- long_main_frames[SOURCE_TYPE == p]
  mat <- dcast(sub[, .N, by = .(destination, dominant_frame)],
                destination ~ dominant_frame,
                value.var = "N", fill = 0L)

  dest_vec <- mat$destination
  mat_m <- as.matrix(mat[, -"destination"])
  rownames(mat_m) <- dest_vec

  # drop destinations with zero counts on all frames (safety)
  keep_rows <- rowSums(mat_m) >= config_2e$min_dest_platform_pairs
  if (sum(keep_rows) < 3) {
    cat("Platform", p,
        "has fewer than 3 destinations with sufficient data; skipping chi-sq.\n")
    next
  }
  mat_m <- mat_m[keep_rows, , drop = FALSE]

  chi <- chisq.test(mat_m)

  n_total <- sum(mat_m)
  k <- min(nrow(mat_m), ncol(mat_m))
  cramers_v <- sqrt(unname(chi$statistic) / (n_total * (k - 1)))

  chi_by_platform[[p]] <- list(
    platform   = p,
    n_pairs    = sum(mat_m),
    n_dest     = nrow(mat_m),
    chi_sq     = unname(chi$statistic),
    df         = chi$parameter,
    p_value    = chi$p.value,
    cramers_v  = cramers_v
  )
}

chi_summary <- rbindlist(lapply(chi_by_platform, as.data.table))
chi_summary[, chi_sq := round(chi_sq, 1)]
chi_summary[, cramers_v := round(cramers_v, 3)]
chi_summary[, significant := p_value < 0.001]

kable(chi_summary, format.args = list(big.mark = ","),
      caption = "Chi squared test of frame by destination, within each platform")

fwrite(chi_summary, "output/tables/02e_chi_by_platform.csv")
```

# 5. Platform Shift Per Destination

For each destination, compare framing profile across platforms. Total
variation distance (TVD) measures how different two probability
distributions are. TVD = 0 means identical profile, TVD = 1 means
completely disjoint. Summary statistic per destination is the maximum
pairwise TVD across platforms.

```{r}
#| label: platform-shift

# destination-platform profile matrix: pct of each frame per (dest, platform)
profile_wide <- dcast(platform_frame_tbl,
                       destination + SOURCE_TYPE ~ dominant_frame,
                       value.var = "pct", fill = 0)

# replace NA with 0 and keep only destinations present on all main platforms
dest_platform_coverage <- profile_wide[, .(n_platforms = .N),
                                         by = destination]
dest_all_platforms <- dest_platform_coverage[n_platforms == length(main_platforms)
                                              ]$destination

profile_valid <- profile_wide[destination %in% dest_all_platforms]

cat("Destinations present on all main platforms:\n")
cat("  ", paste(dest_all_platforms, collapse = ", "), "\n\n")

# compute pairwise TVD for each destination across platforms
tvd_results <- list()

for (d in dest_all_platforms) {
  sub_prof <- profile_valid[destination == d]
  mat_prof <- as.matrix(sub_prof[, c("religious_only",
                                       "religious_tourism_mixed",
                                       "tourism_only")])
  rownames(mat_prof) <- sub_prof$SOURCE_TYPE

  # normalise each row to sum to 1
  mat_prof <- mat_prof / rowSums(mat_prof)

  # pairwise TVD
  n_p <- nrow(mat_prof)
  tvd_mat <- matrix(0, n_p, n_p,
                     dimnames = list(rownames(mat_prof),
                                      rownames(mat_prof)))
  for (i in 1:n_p) {
    for (j in 1:n_p) {
      tvd_mat[i, j] <- 0.5 * sum(abs(mat_prof[i, ] - mat_prof[j, ]))
    }
  }

  tvd_results[[d]] <- list(
    destination     = d,
    tvd_matrix      = tvd_mat,
    mean_tvd        = mean(tvd_mat[upper.tri(tvd_mat)]),
    max_tvd         = max(tvd_mat)
  )
}

tvd_summary <- rbindlist(lapply(tvd_results, function(x)
  data.table(destination = x$destination,
              mean_tvd = round(x$mean_tvd, 3),
              max_tvd = round(x$max_tvd, 3))))
tvd_summary <- tvd_summary[order(-mean_tvd)]

tvd_summary[, stability := fcase(
  max_tvd < 0.10, "stable",
  max_tvd < 0.25, "moderate shift",
  default = "large shift"
)]

kable(tvd_summary,
      caption = "Platform shift per destination (total variation distance)")
fwrite(tvd_summary, "output/tables/02e_platform_shift.csv")
```

# 6. Visualisation

```{r}
#| label: plot-frame-grid

long_main_frames[, destination := factor(destination,
  levels = long_main_frames[, .N,
                             by = destination][order(-N)]$destination)]

plot_data <- long_main_frames[, .N,
                                by = .(SOURCE_TYPE, destination, dominant_frame)]
plot_data[, total := sum(N), by = .(SOURCE_TYPE, destination)]
plot_data[, pct := N / total * 100]

p_grid <- ggplot(plot_data,
                  aes(destination, pct, fill = dominant_frame)) +
  geom_col() +
  facet_wrap(~ SOURCE_TYPE, ncol = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Frame composition by destination and platform",
       x = NULL, y = "Percent of pairs", fill = "Frame") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "bottom")

ggsave("output/figures/02e_frame_grid_by_platform.png", p_grid,
       width = 13, height = 9)
p_grid
```

```{r}
#| label: plot-shift

p_shift <- ggplot(tvd_summary,
                   aes(reorder(destination, mean_tvd),
                       mean_tvd, fill = stability)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("stable" = "#4F8A4F",
                                 "moderate shift" = "#F4A582",
                                 "large shift" = "#B2182B")) +
  geom_hline(yintercept = c(0.10, 0.25),
             linetype = "dashed", colour = "grey50") +
  labs(title = "Platform shift per destination",
       subtitle = "Mean pairwise total variation distance across platforms",
       x = NULL, y = "Mean TVD", fill = "Stability") +
  theme_minimal()

ggsave("output/figures/02e_platform_shift.png", p_shift,
       width = 10, height = 6)
p_shift
```

# 7. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2e\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. at least three platforms met the minimum volume threshold
checks$enough_platforms <- length(main_platforms) >= 3
cat("1. At least 3 platforms with >= ",
    config_2e$min_platform_pairs, " pairs: ",
    ifelse(checks$enough_platforms, "PASS", "FAIL"),
    " (", length(main_platforms), ")\n", sep = "")

# 2. chi squared significant on at least the largest platform
web_chi_sig <- chi_summary[platform == main_platforms[1]]$p_value < 0.001
checks$main_platform_sig <- isTRUE(web_chi_sig)
cat("2. Frame-by-destination significant on largest platform: ",
    ifelse(checks$main_platform_sig, "PASS", "NOTE"), "\n", sep = "")

# 3. at least one destination has mean TVD < 0.15 (stable)
checks$some_stability <- any(tvd_summary$mean_tvd < 0.15)
cat("3. At least one destination is platform stable (TVD < 0.15): ",
    ifelse(checks$some_stability, "PASS", "NOTE"), "\n", sep = "")

# 4. cross-platform TVD spans substantively meaningful range
checks$tvd_range <- diff(range(tvd_summary$mean_tvd)) >= 0.05
cat("4. TVD range across destinations >= 0.05: ",
    ifelse(checks$tvd_range, "PASS", "NOTE"),
    " (range = ", round(diff(range(tvd_summary$mean_tvd)), 3), ")\n", sep = "")

cat("\n")
cat("Passed:", sum(unlist(checks)), "of", length(checks), "\n")
cat(strrep("=", 60), "\n")
```

# 8. Save Outputs

```{r}
#| label: save

meta_2e <- list(
  timestamp        = Sys.time(),
  phase            = "2e_per_platform",
  config           = config_2e,
  platform_counts  = platform_counts,
  main_platforms   = main_platforms,
  chi_by_platform  = chi_summary,
  tvd_summary      = tvd_summary,
  tvd_results      = tvd_results,
  checks           = checks
)
saveRDS(meta_2e, config_2e$output_metadata)
cat("Phase 2e saved to", config_2e$output_metadata, "\n")
```

# 9. Interpretive Summary

Phase 2e provides the platform invariance check for the Phase 2c
framing analysis. The chi squared tests establish whether destination
framing remains statistically structured within each platform
separately. If all platform level chi squared tests are significant,
the framing structure is not a statistical artefact of pooling across
platforms.

The TVD analysis quantifies how much each destination's frame profile
changes across platforms. Low TVD destinations have consistent media
identities regardless of channel. High TVD destinations have platform
specific constructions, suggesting that social media audiences and
portal audiences encounter different versions of the same site.

Both outputs feed the Discursive Framing results section as a robustness
subsection, titled provisionally *Platform Invariance of Framing
Patterns*. The finding that framing structure holds within each
platform strengthens the generalisability of the aggregate results. Any
destination with particularly high TVD deserves narrative attention in
the Discussion, since it reveals platform as an active mediator of
meaning construction rather than a neutral channel.
