---
title: "Phase 2f: Coexposure Network"
subtitle: "Intertextual links between destinations through co-mention graph"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(ggplot2)
library(knitr)
library(igraph)
library(ggraph)
library(tidygraph)

for (d in c("data", "output/tables", "output/figures")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design Notes

Phase 2f analyses the intertextual structure of the destination corpus.
While previous phases treat each destination as an isolated unit of
analysis, articles in the corpus sometimes mention more than one
destination. These co-mentions carry information about how media
construct the broader category of religious tourism in Croatia as a
relational system rather than a set of disconnected sites.

Two analytical components.

First, a coexposure matrix built by taking the article-destination
incidence matrix and computing the pairwise dot product. Each cell
gives the number of articles that mention both destinations. Normalised
using the Jaccard index, cell values range from zero to one and
represent the proportion of shared articles relative to combined
coverage.

Second, a graph representation where nodes are destinations, edges are
weighted by co-mention frequency or Jaccard similarity, and communities
are detected through modularity based clustering. The graph summarises
which destinations form clusters in the media ecosystem and which
remain isolated.

The analysis complements Phase 2c clustering, which groups destinations
by their frame profiles. Phase 2f clusters them by their intertextual
neighbourhood. If the two clusterings align, the media construct
destinations into a small number of tightly coupled groups. If they
diverge, frame similarity and co-mention structure represent
independent relational layers.

# 0. Configuration

```{r}
#| label: config

config_2f <- list(
  timestamp          = Sys.time(),
  input_corpus       = "data/destinations_analytical_cleaned.rds",
  output_metadata    = "data/phase2f_metadata.rds",
  min_edge_weight    = 5L,
  seed               = 2026L
)
set.seed(config_2f$seed)
saveRDS(config_2f, "data/config_phase2f.rds")
cat("Phase 2f configuration saved.\n")
```

# 1. Load Corpus

```{r}
#| label: load

long <- readRDS(config_2f$input_corpus)
setDT(long)
long <- long[cleaning_action == "keep"]

cat("Corpus loaded.\n")
cat("  Total pairs:", format(nrow(long), big.mark = ","), "\n")
cat("  Unique articles:",
    format(uniqueN(long$article_id), big.mark = ","), "\n\n")
```

# 2. Article Level Destination Incidence

```{r}
#| label: incidence

# each row: one article, with one-hot encoded destination membership
art_dest <- unique(long[, .(article_id, destination)])

# count destinations per article
n_dest_per_article <- art_dest[, .N, by = article_id]
setnames(n_dest_per_article, c("article_id", "n_destinations"))

dist_tbl <- n_dest_per_article[, .N, by = n_destinations][order(n_destinations)]
kable(dist_tbl, format.args = list(big.mark = ","),
      caption = "Destinations mentioned per article")

multi_articles <- n_dest_per_article[n_destinations >= 2]$article_id
cat("\nArticles mentioning at least two destinations: ",
    format(length(multi_articles), big.mark = ","), "\n\n", sep = "")
```

# 3. Coexposure Matrix

```{r}
#| label: coexposure-matrix

# incidence matrix (articles x destinations)
incid_long <- unique(long[, .(article_id, destination)])
incid_long[, present := 1L]
incid_wide <- dcast(incid_long, article_id ~ destination,
                     value.var = "present", fill = 0L)

destinations <- setdiff(names(incid_wide), "article_id")
M <- as.matrix(incid_wide[, ..destinations])

# co-occurrence: t(M) %*% M
co_mat <- t(M) %*% M

# diagonal is total article count per destination (self-mentions)
totals <- diag(co_mat)

# zero out diagonal
diag(co_mat) <- 0L

co_dt <- as.data.table(as.table(co_mat))
setnames(co_dt, c("dest_a", "dest_b", "n_shared"))
co_dt <- co_dt[as.character(dest_a) < as.character(dest_b)]
co_dt <- co_dt[n_shared > 0][order(-n_shared)]

kable(head(co_dt, 20), format.args = list(big.mark = ","),
      caption = "Top 20 coexposure pairs (articles mentioning both)")

fwrite(co_dt, "output/tables/02f_coexposure_counts.csv")
```

# 4. Jaccard Similarity

```{r}
#| label: jaccard

# Jaccard for pair (A, B) = |A n B| / |A u B|
# where |A| = n articles mentioning A, |B| similar, |A n B| = n_shared
# |A u B| = |A| + |B| - |A n B|

jaccard_mat <- matrix(0, length(destinations), length(destinations),
                       dimnames = list(destinations, destinations))
for (i in seq_along(destinations)) {
  for (j in seq_along(destinations)) {
    if (i == j) next
    n_a <- totals[i]
    n_b <- totals[j]
    n_shared_ij <- co_mat[i, j]
    union_ij <- n_a + n_b - n_shared_ij
    if (union_ij > 0) {
      jaccard_mat[i, j] <- n_shared_ij / union_ij
    }
  }
}

jaccard_dt <- as.data.table(as.table(jaccard_mat))
setnames(jaccard_dt, c("dest_a", "dest_b", "jaccard"))
jaccard_dt <- jaccard_dt[as.character(dest_a) < as.character(dest_b)]
jaccard_dt[, jaccard := round(jaccard, 4)]
jaccard_dt <- jaccard_dt[jaccard > 0][order(-jaccard)]

kable(head(jaccard_dt, 20),
      caption = "Top 20 Jaccard similarity pairs")

fwrite(jaccard_dt, "output/tables/02f_jaccard.csv")
```

# 5. Graph Construction

```{r}
#| label: graph

# edge list from raw coexposure counts
edges <- co_dt[n_shared >= config_2f$min_edge_weight,
                .(from = as.character(dest_a),
                  to = as.character(dest_b),
                  weight = n_shared)]

# node attributes
nodes <- data.table(name = destinations,
                     total_articles = totals[destinations])

g <- graph_from_data_frame(edges, directed = FALSE,
                             vertices = nodes)

cat("Graph constructed.\n")
cat("  Vertices:", vcount(g), "\n")
cat("  Edges:", ecount(g), "\n")
cat("  Edge weight range:", min(E(g)$weight), "to",
    max(E(g)$weight), "\n\n")
```

# 6. Centrality Measures

```{r}
#| label: centrality

V(g)$degree <- igraph::degree(g)
V(g)$strength <- igraph::strength(g)
V(g)$betweenness <- igraph::betweenness(g, weights = 1 / E(g)$weight)
V(g)$eigen <- eigen_centrality(g, weights = E(g)$weight)$vector

centrality_dt <- data.table(
  destination = V(g)$name,
  total_articles = V(g)$total_articles,
  degree = V(g)$degree,
  strength = V(g)$strength,
  betweenness = round(V(g)$betweenness, 2),
  eigen = round(V(g)$eigen, 3)
)[order(-strength)]

kable(centrality_dt, format.args = list(big.mark = ","),
      caption = "Centrality measures per destination")

fwrite(centrality_dt, "output/tables/02f_centrality.csv")
```

# 7. Community Detection

```{r}
#| label: communities

# Louvain community detection (modularity optimisation)
set.seed(config_2f$seed)
communities <- cluster_louvain(g, weights = E(g)$weight)

V(g)$community <- membership(communities)

modularity_score <- modularity(communities)
cat("Louvain community detection\n")
cat("  Number of communities:", length(communities), "\n")
cat("  Modularity:", round(modularity_score, 3), "\n\n")

community_dt <- data.table(
  destination = V(g)$name,
  community = V(g)$community
)[order(community, destination)]

kable(community_dt, caption = "Community assignment")
fwrite(community_dt, "output/tables/02f_communities.csv")
```

# 8. Graph Visualisation

```{r}
#| label: graph-plot

# convert to tidygraph for ggraph
g_tidy <- as_tbl_graph(g)

set.seed(config_2f$seed)

p_net <- ggraph(g_tidy, layout = "fr", weights = weight) +
  geom_edge_link(aes(width = weight, alpha = weight),
                  colour = "grey50") +
  geom_node_point(aes(size = total_articles,
                        colour = as.factor(community))) +
  geom_node_text(aes(label = name),
                   repel = TRUE, size = 4) +
  scale_edge_width(range = c(0.3, 2.5)) +
  scale_edge_alpha(range = c(0.3, 0.9)) +
  scale_size_continuous(range = c(4, 14)) +
  scale_colour_brewer(palette = "Set2") +
  labs(title = "Destination coexposure network",
       subtitle = paste0("Edge weight = number of articles mentioning both (>=",
                          config_2f$min_edge_weight, ")"),
       size = "Article volume",
       colour = "Community",
       edge_width = "Coexposure",
       edge_alpha = "Coexposure") +
  theme_graph()

ggsave("output/figures/02f_coexposure_network.png", p_net,
       width = 11, height = 9)
p_net
```

```{r}
#| label: jaccard-heatmap

j_plot <- melt(as.data.table(jaccard_mat, keep.rownames = "dest_a"),
                id.vars = "dest_a",
                variable.name = "dest_b",
                value.name = "jaccard")

# order by total articles
dest_order <- centrality_dt[order(-total_articles)]$destination
j_plot[, dest_a := factor(dest_a, levels = dest_order)]
j_plot[, dest_b := factor(dest_b, levels = dest_order)]

p_jacc <- ggplot(j_plot, aes(dest_a, dest_b, fill = jaccard)) +
  geom_tile() +
  geom_text(aes(label = ifelse(jaccard > 0,
                                 format(round(jaccard, 3), nsmall = 3),
                                 "")),
             size = 3) +
  scale_fill_gradient(low = "white", high = "#B2182B") +
  labs(title = "Jaccard similarity between destinations",
       subtitle = "Proportion of shared articles relative to combined coverage",
       x = NULL, y = NULL, fill = "Jaccard") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        panel.grid = element_blank())

ggsave("output/figures/02f_jaccard_heatmap.png", p_jacc,
       width = 10, height = 8)
p_jacc
```

# 9. Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2f\n")
cat(strrep("=", 60), "\n\n")

checks <- list()

# 1. non-trivial number of multi-destination articles
checks$multi_article_threshold <- length(multi_articles) >= 200
cat("1. At least 200 multi-destination articles: ",
    ifelse(checks$multi_article_threshold, "PASS", "NOTE"),
    " (", length(multi_articles), ")\n", sep = "")

# 2. graph is connected (no isolated destinations)
checks$connected <- is_connected(g)
cat("2. Graph is connected (no isolated destinations): ",
    ifelse(checks$connected, "PASS", "NOTE"), "\n", sep = "")

# 3. modularity score indicates meaningful community structure
# > 0.3 is conventionally interpreted as non-trivial structure
checks$modularity_ok <- modularity_score > 0.15
cat("3. Modularity > 0.15 (non-trivial communities): ",
    ifelse(checks$modularity_ok, "PASS", "NOTE"),
    " (", round(modularity_score, 3), ")\n", sep = "")

# 4. at least two communities detected
checks$community_count <- length(communities) >= 2
cat("4. At least 2 communities detected: ",
    ifelse(checks$community_count, "PASS", "NOTE"),
    " (", length(communities), ")\n", sep = "")

cat("\n")
cat("Passed:", sum(unlist(checks)), "of", length(checks), "\n")
cat(strrep("=", 60), "\n")
```

# 10. Save Outputs

```{r}
#| label: save

meta_2f <- list(
  timestamp         = Sys.time(),
  phase             = "2f_coexposure",
  config            = config_2f,
  coexposure_counts = co_dt,
  jaccard           = jaccard_dt,
  centrality        = centrality_dt,
  communities       = community_dt,
  modularity        = modularity_score,
  checks            = checks
)
saveRDS(meta_2f, config_2f$output_metadata)
saveRDS(g, "data/coexposure_graph.rds")
cat("Phase 2f saved to", config_2f$output_metadata, "\n")
```

# 11. Interpretive Summary

Phase 2f reveals the relational structure of the destination corpus.
Three analytical outputs enter the Descriptive Analysis section.

The top coexposure pairs identify destinations that media consistently
construct together. High coexposure between Marija Bistrica and Trsat
would indicate that the national Marian shrine discourse operates as a
unified media category. High coexposure between Nin and Solin would
reveal the starohrvatski historical layer of religious tourism
discourse.

The Jaccard similarity matrix quantifies the relative strength of
co-mention beyond raw frequency. A high Jaccard value means that two
destinations share a large proportion of their coverage relative to
their combined volume, which is a stronger signal of media
co-construction than absolute co-mention counts.

The community detection produces an empirically grounded typology
based on media intertextual structure rather than on destination
content profiles. This typology can be cross referenced with the
Phase 2c framing clusters. Destinations placed in the same community
here and the same framing cluster there represent robust media
categories. Destinations that belong to different communities in the
two analyses reveal a tension between thematic similarity and
intertextual proximity.
