---
title: "Phase 2i: Intercoder Reliability"
subtitle: "Two-pass: generate coding sheets (Pass 1), compute kappa (Pass 2)"
author: "DigiKat Destination Framing Sub-study"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{r}
#| label: setup
#| include: false

library(data.table)
library(stringi)

for (d in c("../../validation/intercoder", "output/tables")) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Design

Two-pass structure. **Pass 1** generates stratified coding sheets and halts.
**Pass 2** runs automatically once both coder sheets are returned and saved to
`validation/intercoder/`.

Valid frame values for coders:
`religious_only | tourism_only | religious_tourism_mixed | memorial_only | religious_memorial | scandal | other`

---

# Pass 1: Generate Coding Sheets

```{r}
#| label: pass1

pass2_ready <- file.exists("../../validation/intercoder/coding_sheet_coder1.csv") &&
              file.exists("../../validation/intercoder/coding_sheet_coder2.csv") &&
              file.exists("../../validation/intercoder/auto_coded.csv")

if (!pass2_ready) {
  long <- readRDS("data/destinations_analytical_final.rds")
  setDT(long)

  set.seed(2026L)
  sample_dt <- long[, .SD[sample(.N, min(20L, .N))], by = destination]

  coding_sheet <- sample_dt[, .(
    article_id   = pair_id,
    destination,
    title        = TITLE,
    snippet      = MENTION_SNIPPET,
    coder_frame  = NA_character_
  )]

  fwrite(coding_sheet,
         "../../validation/intercoder/coding_sheet_coder1.csv",
         bom = TRUE)
  fwrite(coding_sheet,
         "../../validation/intercoder/coding_sheet_coder2.csv",
         bom = TRUE)

  auto_coded <- sample_dt[, .(
    article_id = pair_id,
    destination,
    auto_frame = dominant_frame
  )]
  fwrite(auto_coded, "../../validation/intercoder/auto_coded.csv")

  cat("Pass 1 complete.\n")
  cat("  Sample size:", nrow(coding_sheet), "articles\n")
  cat("  Coding sheets written to validation/intercoder/\n\n")
  cat("Next steps:\n")
  cat("  1. Give coding_sheet_coder1.csv to Coder 1\n")
  cat("  2. Give coding_sheet_coder2.csv to Coder 2\n")
  cat("  3. Ask coders to fill in the coder_frame column\n")
  cat("  4. Return filled sheets to validation/intercoder/\n")
  cat("  5. Re-render this script to run Pass 2\n\n")
  cat("Valid frame values:\n")
  cat("  religious_only | tourism_only | religious_tourism_mixed\n")
  cat("  memorial_only | religious_memorial | scandal | other\n")

  knitr::knit_exit()
}
```

---

# Pass 2: Compute Reliability

```{r}
#| label: pass2-load

# requires irr package: install.packages("irr")
library(irr)

c1   <- fread("../../validation/intercoder/coding_sheet_coder1.csv")
c2   <- fread("../../validation/intercoder/coding_sheet_coder2.csv")
auto <- fread("../../validation/intercoder/auto_coded.csv")

setnames(c1,   "coder_frame", "frame_c1")
setnames(c2,   "coder_frame", "frame_c2")

merged <- merge(
  c1[, .(article_id, frame_c1)],
  c2[, .(article_id, frame_c2)],
  by = "article_id"
)
merged <- merge(merged, auto[, .(article_id, auto_frame)], by = "article_id")

cat("Merged", nrow(merged), "coded articles\n")
```

```{r}
#| label: coder-coder-kappa

kappa_c1c2 <- irr::kappa2(merged[, .(frame_c1, frame_c2)])
cat("Coder 1 vs Coder 2 kappa:", round(kappa_c1c2$value, 3), "\n")
cat("  n =", kappa_c1c2$n, "\n")
```

```{r}
#| label: auto-vs-human-kappa

merged[, frame_human := fifelse(frame_c1 == frame_c2, frame_c1, NA_character_)]
merged_agreed <- merged[!is.na(frame_human)]

cat("Agreed cases:", nrow(merged_agreed), "of", nrow(merged),
    "(", round(nrow(merged_agreed) / nrow(merged) * 100, 1), "%)\n")

kappa_auto <- irr::kappa2(merged_agreed[, .(frame_auto, frame_human)])
cat("Auto vs human kappa (agreed cases):", round(kappa_auto$value, 3), "\n")
```

```{r}
#| label: disagreements

disagree <- merged[frame_c1 != frame_c2, .(
  article_id,
  destination,
  snippet    = stri_sub(snippet, 1, 120),
  frame_c1,
  frame_c2,
  auto_frame
)]
cat("Disagreements:", nrow(disagree), "\n")
fwrite(disagree, "../../validation/intercoder/disagreements.csv", bom = TRUE)
```

```{r}
#| label: save-kappa

kappa_summary <- data.table(
  comparison = c("coder1_vs_coder2", "auto_vs_human"),
  kappa      = c(kappa_c1c2$value, kappa_auto$value),
  n          = c(kappa_c1c2$n, nrow(merged_agreed))
)
print(kappa_summary)
fwrite(kappa_summary, "output/tables/02i_intercoder_kappa.csv")
```

# Decision Gate

```{r}
#| label: decision-gate

cat(strrep("=", 60), "\n")
cat("DECISION GATE 2i\n")
cat(strrep("=", 60), "\n\n")

checks <- list(
  kappa_c1c2_acceptable = kappa_c1c2$value >= 0.6,
  kappa_auto_acceptable = kappa_auto$value >= 0.6,
  agreement_rate_ok     = nrow(merged_agreed) / nrow(merged) >= 0.7,
  kappa_saved           = file.exists("output/tables/02i_intercoder_kappa.csv")
)

for (nm in names(checks)) {
  cat(nm, ":", ifelse(checks[[nm]], "PASS", "REVIEW"), "\n")
}
cat(strrep("=", 60), "\n")
```
